---
title: "Key Findings"
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    toc: false
---

```{r setup, include=FALSE}
# Setup: Load required libraries, configurations, and datasets.
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

# Load configuration file and helper functions
source("R/config.R")
source("R/prep/analysis/helper_functions.R")

# Helper function to highlight numbers
highlight_number <- function(x) {
  paste0("<span style='background-color: yellow; font-weight: bold;'>", x, "</span>")
}

# Load necessary data files for analysis
load(file = paste0(sp_data_path, "/data/analysis/app/ncrp_yearendpop_filtered.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_to_exclude.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_nofilter.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_with_high_missing.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_with_high_missing_race.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_abolished_parole.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/states_use_other_race_eth.rds"))
load(file = paste0(sp_data_path, "/data/analysis/app/which_overall_year.rds"))
```

## Nationally 

**Summary**: 

This section calculates the **average number of years** individuals spent in prison past their parole eligibility year by **race and ethnicity**. The analysis excludes states with unreliable or missing data and states that abolished discretionary parole to ensure results are robust and representative.

**Finding**: 

```{r results='asis'}
# ---------------------------------------------------------------------------- #
# Years Spent in Prison Past Parole Eligibility
# ---------------------------------------------------------------------------- #
# Combine excluded states
combined_excluded_states <- unique(c(
  states_with_high_missing$state,
  states_with_high_missing_race$state,
  states_to_exclude
))

# Remove states that abolished parole
valid_states <- setdiff(
  ncrp_yearendpop_filtered$state, 
  unique(c(combined_excluded_states, states_abolished_parole$state))
)

# Filter NCRP data to valid states only
ncrp_yearendpop_filtered <- fnc_filter_pe_population_criteria(
  data = ncrp_yearendpop_filtered,
  exclude = states_to_exclude,
  dont_filter = states_nofilter
)

# Filter NCRP data to people currently eligible for parole (also past parole eligibility since it's yearend)
ncrp_past_pe <- ncrp_yearendpop_filtered |>
  filter(state %in% valid_states, parelig_status == "Current")

# Average time served past PE for people still in prison by race and ethnicity and state
avg_past_pe_race_by_state <- ncrp_past_pe |>
  filter(race != "Unknown") |>
  filter(
    state %in% states_use_other_race_eth$state |
      (!state %in% states_use_other_race_eth$state &
         race %in% c("White, non-Hispanic", "Hispanic, any race", "Black, non-Hispanic"))
  ) |>
  mutate(years_to_estimated_pey = abs(years_to_estimated_pey)) |>
  group_by(state, race, rptyear) |>
  summarise(
    avg_years_to_estimated_pey = mean(years_to_estimated_pey, na.rm = TRUE),
    total_years_past_pe = sum(years_to_estimated_pey, na.rm = TRUE),
    people = n(),
    .groups = "drop"
  ) |>
  fnc_filter_by_year(which_overall_year)

# Calculate weighted averages
avg_past_pe_race <- avg_past_pe_race_by_state |>
  group_by(race) |>
  summarise(
    avg_years_to_estimated_pey = sum(avg_years_to_estimated_pey * people, na.rm = TRUE) / sum(people, na.rm = TRUE),
    total_years_past_pe = sum(total_years_past_pe, na.rm = TRUE),
    people = sum(people, na.rm = TRUE),
    .groups = "drop"
  )

# Get states used
included_states <- unique(avg_past_pe_race_by_state$state)
states_using_2018_data <- avg_past_pe_race_by_state |>
  filter(rptyear == 2018) |>
  pull(state) |>
  unique()

# Generate Text Outputs
cat("In 2019, ",
    highlight_number(comma(avg_past_pe_race$people[avg_past_pe_race$race == "White, non-Hispanic"])),
    "White, non-Hispanic individuals collectively spent",
    highlight_number(comma(round(avg_past_pe_race$total_years_past_pe[avg_past_pe_race$race == "White, non-Hispanic"], 1))),
    "years in prison past their parole eligibility year across",
    highlight_number(length(included_states)), "states, serving an average of",
    highlight_number(round(avg_past_pe_race$avg_years_to_estimated_pey[avg_past_pe_race$race == "White, non-Hispanic"], 1)),
    "years beyond parole eligibility.<br><br>")

cat("In 2019, ",
    highlight_number(comma(avg_past_pe_race$people[avg_past_pe_race$race == "Black, non-Hispanic"])),
    "Black, non-Hispanic individuals collectively spent",
    highlight_number(comma(round(avg_past_pe_race$total_years_past_pe[avg_past_pe_race$race == "Black, non-Hispanic"], 1))),
    "years in prison past their parole eligibility year across",
    highlight_number(length(included_states)), "states, serving an average of",
    highlight_number(round(avg_past_pe_race$avg_years_to_estimated_pey[avg_past_pe_race$race == "Black, non-Hispanic"], 1)),
    "years beyond parole eligibility.<br><br>")

cat("In 2019, ",
    highlight_number(comma(avg_past_pe_race$people[avg_past_pe_race$race == "Hispanic, any race"])),
    "Hispanic individuals collectively spent",
    highlight_number(comma(round(avg_past_pe_race$total_years_past_pe[avg_past_pe_race$race == "Hispanic, any race"], 1))),
    "years in prison past their parole eligibility year across",
    highlight_number(length(included_states)), "states, serving an average of",
    highlight_number(round(avg_past_pe_race$avg_years_to_estimated_pey[avg_past_pe_race$race == "Hispanic, any race"], 1)),
    "years beyond parole eligibility.<br><br>")
```

```{r results='asis', echo = FALSE}
# Add notes for included states and states with abolished parole
cat("<em style='font-size: 0.9em;'>Note: The following states used 2018 data due to unreliable 2019 data: ",
    paste(states_using_2018_data, collapse = ", "), "</em><br>")

cat("<em style='font-size: 0.9em;'>Included states (", length(included_states), "):", 
    paste(included_states, collapse = ", "), "</em><br><br>")
```

## By Offense

**Summary**: 

This section analyzes the offense types of individuals held past their parole eligibility year, distinguishing between **violent** and **nonviolent offenses**. Nonviolent offenses are further categorized into Property, Drug, and Public Order offenses.

**Finding**: 

```{r results='asis'}
# Categorize offenses into violent vs. nonviolent, keeping "Unknown" offenses for initial analysis
ncrp_past_pe_offense_initial <- ncrp_past_pe |>
  mutate(offense_group = case_when(
    fbi_index %in% c("Murder or Nonnegligent Manslaughter",
                     "Negligent Manslaughter",
                     "Rape or Sexual Assault",
                     "Robbery",
                     "Aggravated or Simple Assault",
                     "Other Violent Offenses") ~ "Violent",
    fbi_index %in% c("Drug", "Public Order", "Property") ~ "Nonviolent",
    TRUE ~ fbi_index))

# Filter out "Unknown" offenses for further analysis
ncrp_past_pe_offense_filtered <- ncrp_past_pe_offense_initial |>
  filter(offense_group != "Unknown") |>
  fnc_filter_by_year(which_overall_year)

# Check if each state has all nonviolent offense types (Property, Drug, Public Order)
states_with_all_offenses <- ncrp_past_pe_offense_filtered |>
  filter(offense_group == "Nonviolent") |>
  group_by(state) |>
  summarise(has_property = any(fbi_index == "Property"),
            has_drug = any(fbi_index == "Drug"),
            has_public_order = any(fbi_index == "Public Order"),
            .groups = "drop") |>
  filter(has_property & has_drug & has_public_order) |>
  pull(state)

# Summarize counts for nonviolent offense subcategories
# Property, Drug, Public Order
nonviolent_summary <- ncrp_past_pe_offense_filtered |>
  filter(offense_group == "Nonviolent") |>
  group_by(fbi_index) |> 
  summarise(total_people = n(), .groups = "drop")

# Extract counts for subcategories
property_offense_count <- nonviolent_summary$total_people[nonviolent_summary$fbi_index == "Property"]
drug_offense_count <- nonviolent_summary$total_people[nonviolent_summary$fbi_index == "Drug"]
public_order_offense_count <- nonviolent_summary$total_people[nonviolent_summary$fbi_index == "Public Order"]

# Total nonviolent count and states with all offenses
nonviolent_offense_count <- sum(nonviolent_summary$total_people)
states_with_nonviolent <- length(states_with_all_offenses)

# Generate output sentence
cat("Over", highlight_number(comma(nonviolent_offense_count)), 
    "people held past their parole eligibility year were convicted of nonviolent offenses in", 
    highlight_number(states_with_nonviolent), "states. Thousands of people who have not committed violent offenses remain incarcerated after their parole eligibility date. This includes approximately",
    highlight_number(comma(property_offense_count)), "people convicted of property offenses,",
    highlight_number(comma(drug_offense_count)), "convicted of drug offenses, and",
    highlight_number(comma(public_order_offense_count)), "convicted of public order offenses.\n\n")
```

```{r results='asis', echo = FALSE}
# Add notes for included states and states with abolished parole
cat("<em style='font-size: 0.9em;'>Note: The following states used 2018 data due to unreliable 2019 data: ",
    paste(states_using_2018_data, collapse = ", "), "</em><br>")

cat("<em style='font-size: 0.9em;'>Included states (", length(included_states), "):", 
    paste(included_states, collapse = ", "), "</em><br><br>")
```




