---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/config.R")

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/robinaparoleeligibility.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_info_by_state.rds"))

# Visualizations (Parole Eligibility Tab)
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_pe_type.rds"))


state_for_report <- "Georgia"

source("R/pull_state_findings.R")
```

<br>
<div id="state-body-section">

```{r}

div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr);",
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "bold-text", style = "text-align: left;", HTML("Parole Board Members")),
  div(id = "bold-text", style = "text-align: left;", HTML("In Prison Past Their<br>Parole Eligibility Date")),
  div(id = "bold-text", style = "text-align: left;", HTML("Pct. of People in Prison<br>Past Their Parole Eligibility Date"))
)
br()
```

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Date

Many people remain incarcerated past their parole eligibility date due to various reasons such as delays in the parole hearing process or unmet parole conditions. 

```{r}
div(id = "visualization", state_stackedbar_pe_type)
```

### Demographics

### Criminal Offense

### Sentence Length

<br>
<div class="divider"></div>

## Missing Data

This section will have information about missing parole eligibility data for the state once new NCRP data is released.





# Prison Population

The size and composition of the prison population have significant implications for state resources and public safety. 

Sentence about change in year-end population from 2010 to 2020.  
Trend graph

<br>
<div class="divider"></div>

## Demographics

An examination of the demographic composition of Georgia's prison population, including age, gender, race, and other relevant factors.  

<br>
<div class="divider"></div>

## Criminal Offenses

An overview of the types of criminal offenses that lead to incarceration, providing insight into the most common crimes and their prevalence.  

<br>
<div class="divider"></div>

## Sentence Lengths

Analysis of the distribution of sentence lengths within the prison population.  


# Prison Releases

Understanding trends in prison releases is crucial for assessing the effectiveness of parole and reentry programs.  

Sentence about change in releases from 2010 to 2020.
Trend graph

## Prison Release Types

<br>
<div class="divider"></div>

## Demographics

A demographic analysis of individuals released from prison, highlighting trends and patterns in reentry.  

<br>
<div class="divider"></div>

## Timing of Release

An exploration of the timing of prison releases, including how long individuals serve before being released.  

<br>
<div class="divider"></div>

## Time Served by Offense Type

Detailed analysis of the average time served by individuals based on their offense type, providing insights into sentencing practices and their outcomes.  















# Disparities

Examining disparities within the prison system is essential for addressing issues of equity and justice. This section focuses on disparities in incarceration and release practices.  

In prison.

## In Prison Past Their Parole Eligibility Date

<br>
<div class="divider"></div>

## Timing of Release

<br>
<div class="divider"></div>

## Timing of Release by Offense Type

<br>
<div class="divider"></div>

## Time Served 

<br>
<div class="divider"></div>






:::

</div>

---
title: "`r state_for_report`"
---
