---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/config.R")

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/carl_state_notes.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_info_by_state.rds"))

# Parole Eligibility Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_pe_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_ped_fbi_index.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_parole_eligibility_sentlgth.rds"))

# Releases Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_line_releases_by_year.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_pie_release_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_agerlse.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_parole_eligibility_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_offense_los.rds"))


# Disparities Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_race_ped_release.rds"))


state_for_report <- "change_me"

source("R/pull_state_findings.R")
```

<div id="state-body-section">

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("Pct. of People in Prison<br>Past Their Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board Members"))
)
br()
```

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Year

`r paste(state_sentence_parole_eligibility_population, collapse = ", ")`

```{r}
div(id = "visualization", state_stackedbar_pe_type)
```

### Demographics

The demographics of people in prison past their parole eligibility year reveals notable proportions among <b>non-Hispanic Black</b> and <b>White people</b>. Gender distribution indicates a predominance of <b>males</b> over females. Age-wise, the majority of individuals fall within the <b>25-34 years</b> and <b>35-44</b> years brackets.

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_parole_eligibility_race),
  div(id = "small-visualization", state_waffle_parole_eligibility_sex),
  div(id = "small-visualization", state_waffle_parole_eligibility_ageyrend),
)
```

### Criminal Offense

The breakdown of criminal offenses of people in prison past their parole eligibility year reveals a varied landscape, with the majority of people incarcerated for <b>aggravated or simple assault (26%) and property crimes (19%)</b>. Other offense categories make up smaller proportions of this population.

```{r}
div(id = "visualization", state_bubble_ped_fbi_index)
```


### Sentence Length

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br>
<div class="divider"></div>

## Missing Data

This section will have information about missing parole eligibility data for the state once new NCRP data is released.

<br>



# Prison Population

## Prison Population

The size and composition of the prison population have significant implications for state resources and public safety. 

Sentence about change in year-end population from 2010 to 2020.  
Trend graph

<br>
<div class="divider"></div>

## Demographics

An examination of the demographic composition of Georgia's prison population, including age, gender, race, and other relevant factors.  

<br>
<div class="divider"></div>

## Criminal Offenses

An overview of the types of criminal offenses that lead to incarceration, providing insight into the most common crimes and their prevalence.  

<br>
<div class="divider"></div>

## Sentence Lengths

Analysis of the distribution of sentence lengths within the prison population.  










# Prison Releases

## Prison Release Trends

From 2010 to 2020, the number of prison releases has declined, dropping from <b>21,717</b> in 2010 to <b>18,298</b> in 2020. This represents a decrease of <b>16%</b>.

```{r}
div(id = "visualization", state_line_releases_by_year)
```
<br>

The proportions of parole-eligible individuals released versus those not released have shown fluctuating trends from 2010 to 2020. In 2020, <b>40%</b> of parole-eligible individuals were released, while <b>60%</b> remained in prison, reflecting ongoing challenges and changes in parole policies and practices.  

```{r}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. In 2020, <b>45%</b> of releases are conditional, requiring ongoing supervision and compliance, while <b>55%</b> are unconditional, allowing individuals to reintegrate without additional constraints.

```{r}
div(id = "visualization", state_pie_release_type)
```


<br>
<div class="divider"></div>

## Demographics of People Released

The demographics of people released from prison in 2020 reveals notable proportions among <b>non-Hispanic Black</b> and <b>White people</b>. Gender distribution indicates a predominance of <b>males</b> over females. Age-wise, the majority of individuals fall within the <b>25-34 years</b> and <b>35-44</b> years brackets. These findings provide insights into the populations transitioning back into the community.

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_releases_race),
  div(id = "small-visualization", state_waffle_releases_sex),
  div(id = "small-visualization", state_waffle_releases_agerlse),
)
```

<br>
<div class="divider"></div>

## Time Served by Offense Type

Between 2010 and 2020, shifts in average time served by individuals for different offense types have been observed. While some categories, like murder and non-negligent manslaughter, have seen a substantial increase in time served, others, like drug-related offenses, have remained relatively stable. 

```{r}
div(id = "visualization", state_lollipop_offense_los)
```














# Disparities

## Incarceration Rates

<br>
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Year

<br>
<div class="divider"></div>

## Years Spent in Prison After Parole Eligibility

ADD TOTAL HERE

```{r}
div(id = "visualization", state_scatter_race_ped_release)
```

<br>
<div class="divider"></div>

## Timing of Release by Offense Type

```{r}
div(id = "visualization", state_bubble_race_ped_release)
```

<br>
<div class="divider"></div>

## Time Served 

<br>
<div class="divider"></div>






:::

</div>

---
title: "`r state_for_report`"
---
