---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/prep/config.R")

state_for_report <- "change_me"

source("R/prep/pull_state_findings_FINAL.R")
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Prison Population Incarcerated Past Their Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board<br>Members"))
)
```

<br>
<div style="font-size: 0.8em; color: black; text-align: center;">
*Only includes people in prison for new offenses and sentence lengths between 1 to 25 years.
</div>

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Their Parole Eligibility

`r paste(state_sentence_pe_type, collapse = ", ")`

```{r state_stackedbar_pe_type}
div(id = "small-visualization", state_stackedbar_pe_type)
```

<div class="transparentdivider"></div>

### Demographics

TBD barcharts of race, ethnicity, age, and sex<br>




<div class="transparentdivider"></div>

### Offense Type

`r paste(state_sentence_parole_eligibility_fbi_index, collapse = ", ")`

```{r state_bar_ped_fbi_index}
div(id = "visualization", state_bar_ped_fbi_index)
```

<div class="transparentdivider"></div>

### Sentence Length

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r state_bar_parole_eligibility_sentlgth}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br><div class="divider"></div>

## State Notes

"What if we change the section from Missing Data to more of a high-level state-specific Methods section? It can include:

1) Total N for the state  
2) How we determined parole eligibility (e.g., what formulas did we use to estimate parole eligibility year)  
3) Missing data patterns"
- Jess S.

<br><br><br>





# Prison Population

## Prison Population Trends

```{r}
div(id = "visualization", state_line_pop_pe_by_year)
```

<br><div class="divider"></div>

## Demographics

The charts below illustrate the demographics of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.  

```{r state_stackedbar_pe_demographics}
div(id = "small-visualization", state_stacked_bar_pe_race)
div(id = "small-visualization", state_stacked_bar_pe_sex)
div(id = "small-visualization", state_stacked_bar_pe_ageyrend)
```

<div class="divider"></div>

## Offense Type

The charts below illustrate the most serious sentenced offenses of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.

```{r state_stacked_bar_pe_fbi_index}
div(id = "visualization", state_stacked_bar_pe_fbi_index)
```

<div class="divider"></div>

## Sentence Length

The charts below illustrate the sentence lengths of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.

```{r state_stacked_bar_pe_sentlgth}
div(id = "small-visualization", state_stacked_bar_pe_sentlgth)
```

<br><br><br>













# Prison Releases

## Prison Release Trends

`r paste(state_sentence_releases, collapse = ", ")`

```{r state_line_releases_by_year}
div(id = "visualization", state_line_releases_by_year)
```
<br>

`r paste(state_sentence_pe_proportion_released, collapse = ", ")`

```{r state_stackedbar_parole_eligibility_release}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. 

```{r state_pie_release_type}
div(id = "visualization", state_pie_release_type)
```

<br>
<div class="divider"></div>

## Demographics

TBD

<br>
<div class="divider"></div>
<br>











:::

</div>

---
title: "`r state_for_report`"
---
