---
title: 'West Virginia Maps'
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    toc: TRUE
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png",
  fig.align = "center"
)

source("R/config.R")

load(file = paste0(config$sp_data_path, "/data/analysis/leaflet_wv_map.rds"))
```

```{r}
# leaflet_wv_map
div(id = "map", leaflet_wv_map) 
```

Upcoming Additions:

-   Prison Locations
-   Parole/Probation Offices
-   Download Features
-   Tables of Info

When/if we have data dashboard information from DHS to avoid hours of manual searching, I'll include: - ER Visits for Drugs - Emergency Response for Suspected Overdoses - SUD Service Locations - Funded Programs for SUD

When we have data:

-   Recidivism Rates
-   Probation/Parole Revocations

Sources:

-   [Map of Services](https://dhhr.wv.gov/office-of-drug-control-policy/help/Pages/default.aspx)
-   [WV DHS Dashboard](https://dhhr.wv.gov/office-of-drug-control-policy/datadashboard/Pages/default.aspx)
-   2021 Census Data
-   WV DOC Annual Reports
