---
title: "The Discretionary Power of Parole" 
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    toc: false
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png"
)

source("R/prep/config.R")

parole_elig_txt <- "Parole eligibility is a critical milestone in the criminal justice system, marking the point at which incarcerated individuals become eligible for conditional release. This eligibility is determined by several factors including the nature of the crime, the length of the sentence, behavior while incarcerated, and state-specific guidelines.<br>"

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))

num_people_current <- sum(parole_eligibility_table$current_count, na.rm = TRUE) 

# State leaders across the United States are grappling with multiple public safety challenges, including shortages in the correctional workforce, limited access to behavioral health services, and the ongoing task of supporting successful reentry for individuals returning from incarceration. Parole release discretion plays a pivotal role in addressing these challenges within 34 state justice systems.
# 
# In 2020, there were 88,000 people in prison past their parole eligibility year, highlighting the critical role that parole boards play in determining the timing of release. The significant number of people who are parole-eligible but have not been released underscores the importance of timely and appropriate parole decisions. Effective parole decision-making is vital not only to managing prison capacity and addressing resource constraints but also to promoting successful reentry, improving public safety outcomes, and ensuring that correctional systems are aligned with best practices in public safety and rehabilitation.
```

<div id="body-section">


```{r}
# Filter the dataframe where current_count is not NA and abolished_parole is "No"
filtered_df <- parole_eligibility_table[!is.na(parole_eligibility_table$current_count) & parole_eligibility_table$abolished_discretionary_parole == "No", ]

# Calculate the total current_count for this filtered data
total_current_count <- sum(filtered_df$current_count, na.rm = TRUE)

# Calculate the national percentage based on filtered_total_pop
national_percentage <- paste0(round(total_current_count / sum(filtered_df$filtered_total_pop, na.rm = TRUE) * 100, 0), "%")

div(
  id = "grid-container",
  style = "
    display: grid; 
    grid-template-columns: repeat(3, 1fr); 
    grid-template-rows: auto 1fr; 
    grid-row-gap: 0px; 
    grid-column-gap: 20px; 
    align-items: end;
  ",
  div(id = "national-highlight", formattable::comma(num_people_current, digits = 0)),
  img(src = "img/pe_1_in_3.png", style = "margin: 0 auto; max-width: 50%; height: auto; text-align: center;"),
  div(id = "national-highlight", "34"),
  div(id = "bold-text", HTML("People* Prison Past Their<br>Parole Eligibility<br><br>")),
  div(id = "bold-text", HTML("1 in X Incarcerated People* Are In Prison Past Their Parole Eligibility")),
  div(id = "bold-text", HTML("States With<br>Discretionary Parole<br><br>"))
)
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   div(id = "national-highlight", formattable::comma(num_people_current, digits = 0)),
#   # div(id = "national-highlight", national_percentage),
#   img(src = "img/pe_1_in_3.png", style = "margin: 0 auto; max-width: 50%; height: auto; text-align: center;"),
#   div(id = "national-highlight", "34"),
#   div(id = "bold-text", HTML("In Prison Past Their<br>Parole Eligibility Year")),
#   div(id = "bold-text", HTML("1 in 3 People Are In Prison Past<br>Their Parole Eligibility Year")),
#   div(id = "bold-text", HTML("States With<br>Discretionary Parole"))
# )
```
</div>
<br>

<div style="font-size: 0.8em; color: black; text-align: center;">
*Only includes people in prison for new offenses and sentence lengths between 1 to 25 years.
</div>

<div class="divider"></div>
<br>
<div id="map-section">
```{r map}
map_percent |> 
    hc_size(height = 690) |> 
    hc_exporting(enabled = TRUE,
                 buttons = list(contextButton = list(align = 'right',
                                                     verticalAlign = 'top',
                                                     y = 0,
                                                     x = -5))) |> 
    hc_chart(marginTop = 0, marginBottom = -50, height = 740)
```
</div>

<div id="body-section">
<div class="divider"></div>
## Parole Eligibility by State

This table provides a comprehensive overview of state-level prison populations*, highlighting the percentage and number of people in prison past their parole eligibility year, alongside details about whether the state has abolished discretionary parole and the number of parole board members.

```{r table}
# Custom function to create data bars with NA handling
data_bars_custom <- function(value) {
  if (is.na(value)) {
    return("-")
  }
  percentage <- percent(value)
  bar <- div(style = list(
    background = "teal",
    width = percentage,
    height = "10px"
  ))
  div(style = list(display = "flex", alignItems = "center"),
      bar, div(style = list(marginLeft = "8px"), percentage))
}

style_cells <- function(value, color) {
  tags$span(style = paste("color:", color, ";"), value)
}

parole_eligibility_table$filtered_total_pop <- as.numeric(parole_eligibility_table$filtered_total_pop)

# https://kcuilla.github.io/reactablefmtr/articles/data_bars.html
# reactable(
#   parole_eligibility_table,
#   pagination = FALSE,
#   searchable = TRUE,
#   defaultSortOrder = "desc",
#   defaultSorted = "current_perc",
#   defaultColDef = colDef(
#     headerStyle = list(fontWeight = "normal"),
#     sortable = TRUE,
#     na = "-" 
#   ),
#   columns = list(
#     state = colDef(name = "State", 
#                    align = "left", 
#                    width = 175,
#                     cell = function(value) {
#         url <- paste0("https://avparoleproject.netlify.app/state_report_", tolower(gsub(" ", "_", value)))
#         tags$a(href = url, target = "_blank", style = "color:black;", value)
#       }),
#     current_perc = colDef(
#       name = "Pct. of People in Prison",
#       align = "left",
#       width = 150,
#       cell = data_bars_custom
#     ),
#     current_count = colDef(name = "Number of People", 
#                            align = "right",
#                            format = colFormat(separators = TRUE)),
#     filtered_total_pop = colDef(name = "Prison Population*", 
#                                 align = "right",
#                                 format = colFormat(separators = TRUE)                           ),
#     abolished_discretionary_parole = colDef(name = "Abolished Discretionary Parole", 
#                                             align = "right"
#                                             ),
#     parole_board_members = colDef(name = "Parole Board Members", 
#                                   align = "right"
#                                   ),
#     ratio = colDef(show = F, name = "Ratio", 
#                    align = "center")
#   ),
#   columnGroups = list(
#     colGroup(name = "In Prison Past Their Parole Eligibility Year", columns = c("current_perc", "current_count"))
#   )
# )
# 
reactable(
  parole_eligibility_table,
  pagination = FALSE,
  searchable = TRUE,
  defaultSortOrder = "desc",
  defaultSorted = "current_perc",
  defaultColDef = colDef(
    headerStyle = list(fontWeight = "normal"),
    sortable = TRUE,
    na = "-" 
  ),
  columns = list(
    state = colDef(
      name = "State", 
      align = "left", 
      width = 175,
      cell = function(value, index) {
        # Use index to reference the correct row
        if (parole_eligibility_table$abolished_discretionary_parole[index] == "No") {
          url <- paste0("https://avparoleproject.netlify.app/state_report_", tolower(gsub(" ", "_", value)))
          tags$a(href = url, target = "_blank", style = "color:black;", value)
        } else {
          return(value)
        }
      }
    ),
    current_perc = colDef(
      name = "Pct. of People in Prison",
      align = "left",
      width = 150,
      cell = data_bars_custom
    ),
    current_count = colDef(
      name = "Number of People", 
      align = "right",
      format = colFormat(separators = TRUE)
    ),
    filtered_total_pop = colDef(
      name = "Prison Population*", 
      align = "right",
      format = colFormat(separators = TRUE)
    ),
    abolished_discretionary_parole = colDef(
      name = "Abolished Discretionary Parole", 
      align = "right"
    ),
    parole_board_members = colDef(
      name = "Parole Board Members", 
      align = "right"
    ),
    ratio = colDef(
      show = F, 
      name = "Ratio", 
      align = "center"
    )
  ),
  columnGroups = list(
    colGroup(
      name = "In Prison Past Their Parole Eligibility Year", 
      columns = c("current_perc", "current_count")
    )
  )
)
```

<div id="footnotes-text"><br>*Includes people in prison for new crimes and with original sentence lengths between 1 to 25 years. Does not include people in prison for parole revocations or sentence lengths less than one year or more than 25 years.<br></div>

<a href="" style="font-size: 0.7em; color: gray;">Download Table</a>

<br>
</div>
<br>







