---
title: "National Snapshot" 
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    toc: false
---

While the most recent available data from the National Corrections Reporting Program (NCRP) is from 2020, we've projected the numbers for 2023 to offer a more current view of how many people are in prison beyond their parole eligibility and how this impacts prison populations.  

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png"
)

source("R/prep/config.R")

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))

# Custom function to round
fnc_round_to_power <- function(x) {
  # Check if the value is NA, and return NA if true
  if (is.na(x)) {
    return(NA)
  }
  
  # Determine the number of digits in the number
  digits <- nchar(floor(x))
  
  # Define the rounding level: if digits >= 3, round to the next power of 10 down, else round to 10
  if (digits >= 3) {
    power <- 10^(digits - 2) # This rounds down to the next power of 10 below the number
    floor(x / power) * power  # Use floor to always round down
  } else {
    round(x, -1)
  }
}

num_people_current <- sum(parole_eligibility_table$current_count, na.rm = TRUE)
num_people_current_rounded <- fnc_round_to_power(num_people_current)
```

<div id="body-section">

```{r highlights}
# Filter the dataframe where current_count is not NA and abolished_parole is "No"
filtered_df <- parole_eligibility_table[!is.na(parole_eligibility_table$current_count) & parole_eligibility_table$abolished_parole== "N", ]

# Calculate the total current_count for this filtered data
total_current_count <- sum(filtered_df$current_count, na.rm = TRUE)

# Calculate the national percentage based on filtered_total_pop
national_percentage <- paste0(round(total_current_count / sum(filtered_df$filtered_total_pop, na.rm = TRUE) * 100, 0), "%")

div(
  id = "grid-container",
  style = "
    display: grid; 
    grid-template-columns: repeat(3, 1fr); 
    grid-template-rows: auto 1fr; 
    grid-row-gap: 0px; 
    grid-column-gap: 20px; 
    align-items: end;
  ",
  div(id = "national-highlight", formattable::comma(num_people_current_rounded, digits = 0)),
  img(src = "img/pe_1_in_3.png", style = "margin: 0 auto; max-width: 50%; height: auto; text-align: center;"),
  div(id = "national-highlight", "34"),
  div(id = "bold-text", HTML("People In Prison Past Their<br>Parole Eligibility*<br><br>")),
  div(id = "bold-text", HTML("1 in X Incarcerated People Are In Prison Past Their Parole Eligibility")),
  div(id = "bold-text", HTML("States With<br>Discretionary Parole**<br><br>"))
)
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   div(id = "national-highlight", formattable::comma(num_people_current, digits = 0)),
#   # div(id = "national-highlight", national_percentage),
#   img(src = "img/pe_1_in_3.png", style = "margin: 0 auto; max-width: 50%; height: auto; text-align: center;"),
#   div(id = "national-highlight", "34"),
#   div(id = "bold-text", HTML("In Prison Past Their<br>Parole Eligibility Year")),
#   div(id = "bold-text", HTML("1 in 3 People Are In Prison Past<br>Their Parole Eligibility Year")),
#   div(id = "bold-text", HTML("States With<br>Discretionary Parole"))
# )
```
</div>
<br>
<div style="font-size: 0.7em; color: black; text-align: left;">
<p>*Only includes people in prison for new offenses and excludes people with life sentences and sentences less than one year. 
**Louisiana is listed among the states with parole systems, despite its recent abolition of parole, due to a substantial population that remains eligible for parole release under the previous system.
</div>
<br>
<div class="divider"></div>
<br>
<div id="map-section">
```{r map}
map_percent |>
    hc_exporting(enabled = TRUE,
                 buttons = list(contextButton = list(align = 'right',
                                                     verticalAlign = 'top',
                                                     y = 0,
                                                     x = -5)))
```
</div>
<br>
<div id="body-section">
<div class="divider"></div>

## Parole Eligibility Data by State

This table provides a detailed projection of the 2023 snapshot of state-level prison populations, including the percentage and number of people who are incarcerated beyond their parole eligibility. It also highlights the number of parole board members for each state.

```{r table}
# Custom function to create data bars with NA handling
data_bars_custom <- function(value) {
  if (is.na(value)) {
    return("-")
  }
  percentage <- percent(value)
  bar <- div(style = list(
    background = "teal",
    width = percentage,
    height = "10px"
  ))
  div(style = list(display = "flex", alignItems = "center"), bar,
      div(style = list(marginLeft = "5px"), percentage)
      )
}

style_cells <- function(value, color) {
  tags$span(style = paste("color:", color, ";"), value)
}

parole_eligibility_table$filtered_total_pop <- as.numeric(parole_eligibility_table$filtered_total_pop)

reactable(
  parole_eligibility_table |> select(-current_count_rounded),
  pagination = FALSE,
  searchable = TRUE,
  defaultSorted = c("abolished_parole", "current_perc"),
  defaultColDef = colDef(
    headerStyle = list(fontWeight = "normal"),
    sortable = TRUE,
    na = "-" 
  ),
  columns = list(
    state = colDef(
      name = "State", 
      align = "left", 
      width = 175,
      cell = function(value, index) {
        # Check if current_perc is NA or if discretionary parole is abolished
        if (!is.na(parole_eligibility_table$current_perc[index])) {
          url <- paste0("https://avparoleproject.netlify.app/state_report_", tolower(gsub(" ", "_", value)))
          tags$a(href = url, target = "_blank", style = "color:black;", value)
        } else {
          return(value)
        }
      }
    ),
    current_perc = colDef(
      name = "Percentage of People in Prison",
      align = "left",
      width = 250,
      cell = data_bars_custom,
      defaultSortOrder = "desc"  # Sort descending
    ),
    current_count = colDef(
      name = "Number of People", 
      align = "right",
      format = colFormat(separators = TRUE),
       cell = function(value, index) {
          # Check if the value is NA (no data)
          if (is.na(value)) {
            return("-")
          }
          # Otherwise, round the value using the custom function and add commas
          else {
            rounded_value <- fnc_round_to_power(value)
            formatted_value <- format(rounded_value, big.mark = ",", scientific = FALSE)
            return(formatted_value)
          }
        }
      ),
    filtered_total_pop = colDef(
      name = "Prison Population", 
      align = "right",
      format = colFormat(separators = TRUE)
    ),
    abolished_parole = colDef(
      show = F,
      name = "Abolished Discretionary Parole", 
      align = "right",
      width = 150,
      defaultSortOrder = "asc"  # Sort ascending
    ),
    members = colDef(
      name = "Parole Board Members*", 
      align = "right"
    ),
    ratio = colDef(
      show = F, 
      name = "Ratio", 
      align = "center"
    )
  ),
  columnGroups = list(
    colGroup(
      name = "People In Prison Past Parole Eligibility", 
      columns = c("current_perc", "current_count")
    )
  )
)
```

<div id="footnotes-text"><br>*The number of parole board members includes appointed members and does not include subordinate hearing examiners that many states employ.<br></div>
<a href="" style="font-size: 0.7em; color: gray;">Download Table</a>
<br>
</div>
<div class="divider"></div>
<br>

