---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/prep/config.R")

state_for_report <- "Georgia"

source("R/prep/pull_state_findings_FINAL.R")
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Prison Population Incarcerated Past Their Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board<br>Members"))
)
```

<br>
<div style="font-size: 0.8em; color: black; text-align: center;">
*Only includes people in prison for new offenses and sentence lengths between 1 to 25 years.
</div>

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Their Parole Eligibility

`r paste(state_sentence_pe_type, collapse = ", ")`

```{r state_stackedbar_pe_type}
div(id = "small-visualization", state_stackedbar_pe_type)
```

<div class="transparentdivider"></div>

### Demographics

TBD barcharts of race, ethnicity, age, and sex<br>




<div class="transparentdivider"></div>

### Offense Type

`r paste(state_sentence_parole_eligibility_fbi_index, collapse = ", ")`

```{r state_bar_ped_fbi_index}
div(id = "visualization", state_bar_ped_fbi_index)
```

<div class="transparentdivider"></div>

### Sentence Length

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r state_bar_parole_eligibility_sentlgth}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br><div class="divider"></div>

## State Notes

"What if we change the section from Missing Data to more of a high-level state-specific Methods section? It can include:

1) Total N for the state  
2) How we determined parole eligibility (e.g., what formulas did we use to estimate parole eligibility year)  
3) Missing data patterns"
- Jess S.

<br><br><br>





# Prison Population

## Prison Population Trends

```{r}
div(id = "visualization", state_line_pop_pe_by_year)
```

<br><div class="divider"></div>

## Demographics

The charts below illustrate the demographics of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.  

```{r state_stackedbar_pe_demographics}
div(id = "small-visualization", state_stacked_bar_pe_race)
div(id = "small-visualization", state_stacked_bar_pe_sex)
div(id = "small-visualization", state_stacked_bar_pe_ageyrend)
```

<div class="divider"></div>

## Offense Type

The charts below illustrate the most serious sentenced offenses of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.

```{r state_stacked_bar_pe_fbi_index}
div(id = "visualization", state_stacked_bar_pe_fbi_index)
```

<div class="divider"></div>

## Sentence Length

The charts below illustrate the sentence lengths of people* in prison based on their parole eligibility status, categorized as currently eligible for parole, eligible in the future, and those with missing parole eligibility information.

```{r state_stacked_bar_pe_sentlgth}
div(id = "small-visualization", state_stacked_bar_pe_sentlgth)
```

<br><br><br>













# Prison Releases

## Prison Release Trends

`r paste(state_sentence_releases, collapse = ", ")`

```{r state_line_releases_by_year}
div(id = "visualization", state_line_releases_by_year)
```
<br>

`r paste(state_sentence_pe_proportion_released, collapse = ", ")`

```{r state_stackedbar_parole_eligibility_release}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. 

```{r state_pie_release_type}
div(id = "visualization", state_pie_release_type)
```

<br>
<div class="divider"></div>

## Demographics

TBD

<br>
<div class="divider"></div>

## Time-Served by Offense

Feedback: Not useful, potentially show  "Average Expected Parole Release Year" versus "Average Actual Time Served."
<br>
















# Disparities

## Incarceration Rates

`r paste(state_sentence_rri, collapse = ", ")`

<br>
<div class="divider"></div>

## Years Spent in Prison After Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r state_parole_release_disparities, results='asis'}
df1 <- all_parole_release_disparities |>
  filter(state == state_for_report) |> 
  filter(!is.na(race)) |> 
  select(time_between_ped_release, race) |> 
  group_by(race) |> 
  summarise(total_years = sum(time_between_ped_release, na.rm = TRUE))

# Check if all total_years are NA
if (!all(is.na(df1$total_years))) {
  
  message <- HTML(paste0("In ", select_year, ", people in prison past their parole eligibility year collectively spent a total of ",
                   comma(sum(df1$total_years, na.rm = TRUE)), " years behind bars instead of supervised in the community on parole."))
  
  cat(message)
  
  br()
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Black, non-Hispanic"]) ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "White, non-Hispanic"]) ),
    div(style = "color: #d97d68; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Hispanic, any race"]) ),
    div(style = "color: #decf64; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Other race(s), non-Hispanic"]) ),
    div(style = "color: #49a7a1; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #55b4e5; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #d97d68; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #decf64; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Black People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("White People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Hispanic People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Other Race(s), Non-Hispanic People"))
  )
  
  cat(as.character("<br><br>"))
  cat(as.character(html_output))
  br()
} else {
  cat("")
}
```

### Gender

In progress. Same type of graph as race and ethnicity tab.

:::

<br>
<div class="divider"></div>




## Timing of Release After Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

The timing of release for individuals after their parole eligibility year varied by offense type and demographic group, with many remaining in prison for several years beyond their eligibility date. 

### Gender

In progress. Same type of graph as race and ethnicity tab.

:::

<br>
<div class="divider"></div>




## Time Served

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r}
# section_text <- "X individuals faced the longest average time served in prison in 2020. Similarly, Y, individuals also experienced notably longer prison stays compared to their Z counterparts, who had the shortest average time served."
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   section_text,
#   div(id = "visualization", state_lollipop_los_race)
# )
# br()
```

`r paste(state_sentence_los_race, collapse = ", ")`

```{r state_lollipop_los_race}
div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_los_race)
# div(id = "visualization", style = "margin: 0 auto; text-align: center;", state_lollipop_los_race)
```

`r paste(state_sentence_los_race_offense, collapse = ", ")`

```{r state_scatter_los_race_offense}
div(id = "visualization", state_scatter_los_race_offense)
```

### Gender 

In progress. Same type of graph as race and ethnicity tab.

:::
<br>
<br>












:::

</div>

---
title: "`r state_for_report`"
---
