---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/prep/config.R")

state_for_report <- "Idaho"

source("R/prep/pull_state_findings.R")

rri_infographic_black <- paste0(config$sp_data_path, "/data/analysis/app/rri_infographic_black_", state_for_report, ".png")
rri_infographic_hispanic <- paste0(config$sp_data_path, "/data/analysis/app/rri_infographic_hispanic_", state_for_report, ".png")
pe_rri_infographic_black <- paste0(config$sp_data_path, "/data/analysis/app/pe_rri_infographic_black_", state_for_report, ".png")
pe_rri_infographic_hispanic <- paste0(config$sp_data_path, "/data/analysis/app/pe_rri_infographic_hispanic_", state_for_report, ".png")
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Prison Population Incarcerated Past Their Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board<br>Members"))
)
```

<br>
<div style="font-size: 0.8em; color: black; text-align: center;">
*Only includes people in prison for new offenses and sentence lengths between 1 to 25 years.
</div>

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

Parole eligibility refers to the point in a person's prison sentence when they become eligible for consideration for release under parole. Parole is a conditional release, allowing the person to serve the remainder of their sentence in the community under supervision, rather than in prison.  

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Their Parole Eligibility

`r paste(state_sentence_pe_type, collapse = ", ")`

```{r state_stackedbar_pe_type}
div(id = "small-visualization", state_stackedbar_pe_type)
```

`r paste(state_sentence_pop_pe_by_year, collapse = ", ")`

```{r}
div(id = "visualization", state_stackedbar_pop_pe_by_year)
```

<br><div class="transparentdivider"></div>

### Demographics

::: {.panel-tabset .custom-tabs}

#### Race and Ethnicity

`r paste(state_sentence_parole_eligibility_race, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_parole_eligibility_race)
```

#### Gender

`r paste(state_sentence_parole_eligibility_sex, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_parole_eligibility_sex)
```

#### Age

`r paste(state_sentence_parole_eligibility_ageyrend, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_parole_eligibility_ageyrend)
```

:::

<br><div class="transparentdivider"></div>

### Offense Types

`r paste(state_sentence_parole_eligibility_fbi_index, collapse = ", ")`

```{r state_bar_ped_fbi_index}
div(id = "visualization", state_bar_ped_fbi_index)
```

<br><div class="transparentdivider"></div>

### Sentence Lengths

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r state_bar_parole_eligibility_sentlgth}
div(id = "small-visualization", state_bar_parole_eligibility_sentlgth)
```

<br><div class="divider"></div>

## State Notes

"What if we change the section from Missing Data to more of a high-level state-specific Methods section? It can include:

1) Total N for the state  
2) How we determined parole eligibility (e.g., what formulas did we use to estimate parole eligibility year)  
3) Missing data patterns"
- Jess S.





# Prison Population

## Prison Population Trends

`r paste(state_sentence_population, collapse = ", ")`

```{r state_line_population_by_year}
div(id = "visualization", state_line_population_by_year)
# <a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners Series</a>
```

<br><div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

#### Race and Ethnicity

`r paste(state_sentence_population_race, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_population_race)
```

#### Gender

`r paste(state_sentence_population_sex, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_population_sex)
```

#### Age

`r paste(state_sentence_population_ageyrend, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_population_ageyrend)
```

:::

<br><div class="divider"></div>

## Offense Types

`r paste(state_sentence_population_fbi_index, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_population_fbi_index)
```

<br><div class="divider"></div>

## Sentence Lengths

`r paste(state_sentence_population_sentlgth, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_population_sentlgth)
```













# Prison Releases

## Prison Release Trends

`r paste(state_sentence_releases, collapse = ", ")`

```{r state_line_releases_by_year}
div(id = "visualization", state_line_releases_by_year)
```
<br>

`r paste(state_sentence_pe_proportion_released, collapse = ", ")`

```{r state_stackedbar_parole_eligibility_release}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. 

```{r state_pie_release_type}
div(id = "visualization", state_pie_release_type)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

#### Race and Ethnicity

`r paste(state_sentence_releases_race, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_releases_race)
```
  
#### Gender
  
`r paste(state_sentence_releases_sex, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_releases_sex)
```
  
#### Age
  
`r paste(state_sentence_releases_agerlse, collapse = ", ")`

```{r}
div(id = "small-visualization", state_bar_releases_agerlse)
```

:::

<br>
<div class="divider"></div>

## Offense Types

`r paste(state_sentence_releases_fbi_index, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_releases_fbi_index)
```










# Disparities

## Disparities in Incarceration

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

Racial and ethnic disparities in prison incarceration represents the accumulation of disparities across the system.  
<br>

```{r}
if (file.exists(rri_infographic_black)) {
  knitr::include_graphics(rri_infographic_black)
}
# if (file.exists(rri_infographic_black)) {
#   knitr::asis_output('<span class="black-non-hispanic">Black, Non-Hispanic</span>')
# }
```

`r paste(state_sentence_rri_black, collapse = ", ")`

```{r}
if (file.exists(rri_infographic_hispanic)) {
    knitr::include_graphics(rri_infographic_hispanic)
}
# if (file.exists(rri_infographic_hispanic)) {
#   knitr::asis_output('<span class="hispanic">Hispanic, Any Race</span>')
# }
```

`r paste(state_sentence_rri_hispanic, collapse = ", ")`

### Gender

Gender disparities in prison incarceration represents the accumulation of disparities across the system.  

<br>

:::

<br>
<div class="divider"></div>



## Time Served

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r}
# section_text <- "X individuals faced the longest average time served in prison in 2020. Similarly, Y, individuals also experienced notably longer prison stays compared to their Z counterparts, who had the shortest average time served."
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   section_text,
#   div(id = "visualization", state_lollipop_los_race)
# )
# br()
```

`r paste(state_sentence_los_race, collapse = ", ")`

```{r state_lollipop_los_race}
div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_los_race)
# div(id = "visualization", style = "margin: 0 auto; text-align: center;", state_lollipop_los_race)
```

This chart shows the average time served in prison for the most serious sentenced offense by race and ethnicity. 
`r paste(state_sentence_los_race_offense, collapse = ", ")`

```{r state_scatter_los_race_offense}
div(id = "visualization", state_scatter_los_race_offense)
```

### Gender 

This chart shows the average time served in prison for the most serious sentenced offense by gender. 
`r paste(state_sentence_los_sex, collapse = ", ")`

```{r state_lollipop_los_sex}
div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_los_sex)
# div(id = "visualization", style = "margin: 0 auto; text-align: center;", state_lollipop_los_sex)
```

`r paste(state_sentence_los_sex_offense, collapse = ", ")`

```{r state_scatter_los_sex_offense}
div(id = "visualization", state_scatter_los_sex_offense)
```

:::

<div class="divider"></div>


## Disparities in Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

<br>

```{r}
if (file.exists(pe_rri_infographic_black)) {
  knitr::include_graphics(pe_rri_infographic_black)
}
```

`r paste(state_sentence_pe_rri_black, collapse = ", ")`

```{r}
if (file.exists(pe_rri_infographic_hispanic)) {
    knitr::include_graphics(pe_rri_infographic_hispanic)
}
```

`r paste(state_sentence_pe_rri_hispanic, collapse = ", ")`

### Gender

:::

<br>
<div class="divider"></div>








## Years in Prison After Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

`r paste(state_sentence_avg_pe_release_race, collapse = ", ")`

```{r state_parole_release_disparities, results='asis'}
# df1 <- all_pe_release_total_years_race |>
#   filter(state == state_for_report)
# 
# # Check if all total_years are NA
# if (!all(is.na(df1$total_years))) {
# 
#   message <- HTML(paste0("In ", select_year, ", people in prison collectively spent a total of ",
#                    comma(sum(df1$total_years, na.rm = TRUE)), " years past their parole eligibility."))
# 
#   cat(message)
# 
#   br()
#   html_output <- div(
#     id = "grid-container",
#     style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
#     div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Black, non-Hispanic"]) ),
#     div(style = "color: #decf64; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Hispanic, any race"]) ),
#     div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "White, non-Hispanic"]) ),
#     div(style = "color: #938ebf; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Other race(s), non-Hispanic"]) ),
# 
#     div(style = "color: #49a7a1; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
#     div(style = "color: #decf64; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
#     div(style = "color: #55b4e5; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
#     div(style = "color: #938ebf; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
# 
#     div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Black People")),
#     div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Hispanic People")),
#     div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("White People")),
#     div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Other Race(s), Non-Hispanic People"))
#   )
# 
#   cat(as.character("<br><br>"))
#   cat(as.character(html_output))
#   br()
# } else {
#   cat("")
# }
div(id = "bold-text", "Avg. Years in Prison After Parole Eligibility")
br()

df1 <- ncrp_avg_pe_release_race |>
  filter(state == state_for_report)

# Check if all average_avg_pe_release are NA
if (!all(is.na(df1$average_avg_pe_release))) {
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;", 
        paste0(round(df1$average_avg_pe_release[df1$race == "Black, non-Hispanic"], 1), " Years")
        ),
    div(style = "color: #decf64; text-align: center; font-size: 24px; font-weight: bold;", 
        paste0(round(df1$average_avg_pe_release[df1$race == "Hispanic, any race"], 1), " Years")
        ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;", 
        paste0(round(df1$average_avg_pe_release[df1$race == "White, non-Hispanic"], 1), " Years")
        ),
    div(style = "color: #938ebf; text-align: center; font-size: 24px; font-weight: bold;", 
        paste0(round(df1$average_avg_pe_release[df1$race == "Other race(s), non-Hispanic"], 1), " Years")
        ),

    div(style = "color: #49a7a1; text-align: center; font-size: 16px; font-weight: bold;", HTML("Black")),
    div(style = "color: #decf64; text-align: center; font-size: 16px; font-weight: bold;", HTML("Hispanic")),
    div(style = "color: #55b4e5; text-align: center; font-size: 16px; font-weight: bold;", HTML("White")),
    div(style = "color: #938ebf; text-align: center; font-size: 16px; font-weight: bold;", HTML("Other"))
  )
  cat(as.character(html_output))
  br()
} else {
  cat("")
}
```

```{r state_lollipop_avg_pe_release_race}
# div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_avg_pe_release_race)
```

`r paste(state_sentence_avg_pe_release_race_offense, collapse = ", ")`

```{r state_scatter_avg_pe_release_race_offense}
div(id = "visualization", state_scatter_avg_pe_release_race_offense)
```

### Gender

`r paste(state_sentence_avg_pe_release_sex, collapse = ", ")`

```{r state_pe_release_total_years_sex, results='asis'}
div(id = "bold-text", "Avg. Years in Prison After Parole Eligibility")
br()

df1 <- ncrp_avg_pe_release_sex |>
  filter(state == state_for_report)

# Check if all average_avg_pe_release are NA
if (!all(is.na(df1$average_avg_pe_release))) {
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold; margin: 0; padding: 0;", 
        paste0(round(df1$average_avg_pe_release[df1$sex == "Male"], 1), " Years")
    ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold; margin: 0; padding: 0;", 
        paste0(round(df1$average_avg_pe_release[df1$sex == "Female"], 1), " Years")
    ),
    
    div(style = "color: #49a7a1; text-align: center; font-size: 16px; font-weight: bold; margin: 0; padding: 0;", HTML("Male")),
    div(style = "color: #55b4e5; text-align: center; font-size: 16px; font-weight: bold; margin: 0; padding: 0;", HTML("Female"))
  )
  cat(as.character(html_output))
  br()
} else {
  cat("")
}
```

`r paste(state_sentence_avg_pe_release_sex_offense, collapse = ", ")`

```{r state_scatter_avg_pe_release_sex_offense}
div(id = "visualization", state_scatter_avg_pe_release_sex_offense)
```

:::

<br>
<div class="divider"></div>


## How Disparities Are Calculated

While White, Black, and Hispanic are not the only race and ethnicity categories found in the U.S., NCRP only provides four categories: White, Black, Hispanic, and Other. Further categories will be added in future data collections by the CSG Justice Center. Additionally, the term Hispanic is used here to maintain consistency with the data source material.

State data broken down by race and ethnicity were pulled from National Corrections Reporting Program, 1991-2020: Selected Variables (NCRP). Then, the rate of people readmitted to parole was calculated within each racial and ethnic group based on the Census Public Use Microdata Sample (PUMS). The relative rate index is calculated by dividing the rate for Black individuals or Hispanic individuals by the rate for White individuals.  

*States vary in how they collect and report information about race and ethnicity. Some states use the race or ethnicity recorded by intake officers, while other states allow individuals to identify their own race and ethnicity. Furthermore, some states combine race and ethnicity into one category, while others separate them into two distinct items. These differences can result in an incomplete count of Hispanic individuals and those of mixed race in state reports. The US Census allows individuals to self-identify their race and ethnicity separately. The variation in practices across states and between state and Federal data reporting systems may result in a deflated RRI for Hispanic individuals. These differences can result in an incomplete count of Hispanic individuals and those of mixed race in state reports. Conversely, it can result in an overcount of people of other races, especially non-Hispanic White people.*

:::

<div class="divider"></div>

## End Notes

Data analysis was conducted using information from the National Corrections Reporting Program and the National Prisoner Statistics Program. See “National Corrections Reporting Program (NCRP),” Bureau of Justice Statistics, accessed September 25, 2024, https://bjs.ojp.gov/data-collection/national-corrections-reporting-program-ncrp; “National Prisoner Statistics (NPS) Program,” Bureau of Justice Statistics, accessed September 25, 2024, https://bjs.ojp.gov/data-collection/national-prisoner-statistics-nps-program.  

</div>

---
title: "`r state_for_report`"
---
