---
title: "State Reports"
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png"
)

#######################################
# This code dynamically generates a list of clickable state report links
# based on the states present in the 'parole_eligibility_table' dataframe.
#
# 1. It filters the 'parole_eligibility_table' dataframe to include only
#    states where the 'current_perc' column is not missing.
# 2. It defines a list of URLs for each state report.
# 3. It checks if each state in the list exists in the filtered dataframe.
# 4. If a state exists, it generates an HTML link for that state report.
# 5. The generated HTML links are combined into a single string and
#    output using 'asis_output()' to ensure the links are rendered correctly
#    in the final HTML document.
#######################################

source("R/config.R")

load(file = paste0(sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
states_to_include <- parole_eligibility_table |> filter(!is.na(current_perc))
```

```{r}


# List of states with their corresponding URLs
state_links <- list(
  "Alabama" = "https://avparoleproject.netlify.app/state_report_alabama",
  "Alaska" = "https://avparoleproject.netlify.app/state_report_alaska",
  "Arizona" = "https://avparoleproject.netlify.app/state_report_arizona",
  "Arkansas" = "https://avparoleproject.netlify.app/state_report_arkansas",
  "California" = "https://avparoleproject.netlify.app/state_report_california",
  "Colorado" = "https://avparoleproject.netlify.app/state_report_colorado",
  "Connecticut" = "https://avparoleproject.netlify.app/state_report_connecticut",
  "Delaware" = "https://avparoleproject.netlify.app/state_report_delaware",
  "Florida" = "https://avparoleproject.netlify.app/state_report_florida",
  "Georgia" = "https://avparoleproject.netlify.app/state_report_georgia",
  "Hawaii" = "https://avparoleproject.netlify.app/state_report_hawaii",
  "Idaho" = "https://avparoleproject.netlify.app/state_report_idaho",
  "Illinois" = "https://avparoleproject.netlify.app/state_report_illinois",
  "Indiana" = "https://avparoleproject.netlify.app/state_report_indiana",
  "Iowa" = "https://avparoleproject.netlify.app/state_report_iowa",
  "Kansas" = "https://avparoleproject.netlify.app/state_report_kansas",
  "Kentucky" = "https://avparoleproject.netlify.app/state_report_kentucky",
  "Louisiana" = "https://avparoleproject.netlify.app/state_report_louisiana",
  "Maine" = "https://avparoleproject.netlify.app/state_report_maine",
  "Maryland" = "https://avparoleproject.netlify.app/state_report_maryland",
  "Massachusetts" = "https://avparoleproject.netlify.app/state_report_massachusetts",
  "Michigan" = "https://avparoleproject.netlify.app/state_report_michigan",
  "Minnesota" = "https://avparoleproject.netlify.app/state_report_minnesota",
  "Mississippi" = "https://avparoleproject.netlify.app/state_report_mississippi",
  "Missouri" = "https://avparoleproject.netlify.app/state_report_missouri",
  "Montana" = "https://avparoleproject.netlify.app/state_report_montana",
  "Nebraska" = "https://avparoleproject.netlify.app/state_report_nebraska",
  "Nevada" = "https://avparoleproject.netlify.app/state_report_nevada",
  "New Hampshire" = "https://avparoleproject.netlify.app/state_report_new_hampshire",
  "New Jersey" = "https://avparoleproject.netlify.app/state_report_new_jersey",
  "New Mexico" = "https://avparoleproject.netlify.app/state_report_new_mexico",
  "New York" = "https://avparoleproject.netlify.app/state_report_new_york",
  "North Carolina" = "https://avparoleproject.netlify.app/state_report_north_carolina",
  "North Dakota" = "https://avparoleproject.netlify.app/state_report_north_dakota",
  "Ohio" = "https://avparoleproject.netlify.app/state_report_ohio",
  "Oklahoma" = "https://avparoleproject.netlify.app/state_report_oklahoma",
  "Oregon" = "https://avparoleproject.netlify.app/state_report_oregon",
  "Pennsylvania" = "https://avparoleproject.netlify.app/state_report_pennsylvania",
  "Rhode Island" = "https://avparoleproject.netlify.app/state_report_rhode_island",
  "South Carolina" = "https://avparoleproject.netlify.app/state_report_south_carolina",
  "South Dakota" = "https://avparoleproject.netlify.app/state_report_south_dakota",
  "Tennessee" = "https://avparoleproject.netlify.app/state_report_tennessee",
  "Texas" = "https://avparoleproject.netlify.app/state_report_texas",
  "Utah" = "https://avparoleproject.netlify.app/state_report_utah",
  "Vermont" = "https://avparoleproject.netlify.app/state_report_vermont",
  "Virginia" = "https://avparoleproject.netlify.app/state_report_virginia",
  "Washington" = "https://avparoleproject.netlify.app/state_report_washington",
  "West Virginia" = "https://avparoleproject.netlify.app/state_report_west_virginia",
  "Wisconsin" = "https://avparoleproject.netlify.app/state_report_wisconsin",
  "Wyoming" = "https://avparoleproject.netlify.app/state_report_wyoming"
)

# Create a vector of states present in the dataframe
available_states <- unique(states_to_include$state)

# Generate the HTML links for the states present in the dataframe
html_links <- lapply(names(state_links), function(state) {
  if (state %in% available_states) {
    paste0("[", state, "](", state_links[[state]], ")<br>")
  }
})

# Collapse the list into a single string
html_output <- paste(unlist(html_links), collapse = "")

# Output the HTML using asis_output to render correctly 
knitr::asis_output(html_output)
```

<br><br>

**Missing Data**

The following states are not included due to incomplete data submitted to the National Corrections Reporting Program in 2020, though they still have parole systems:

- Alaska
- Vermont
- Arkansas
- Hawaii

Additionally, the analysis excludes states that have abolished parole, meaning no state reports are provided for:

- Arizona
- California
- Delaware
- Florida
- Illinois
- Indiana
- Kansas
- Maine
- Minnesota
- New Mexico
- North Carolina
- Ohio
- Oregon
- Virginia
- Washington
- Wisconsin


