---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

# load packages
library(shiny)
library(highcharter)
library(dplyr)
library(htmlwidgets)
library(dplyr)
library(htmltools)
library(reactable)
library(reactablefmtr)
library(htmltools)
library(tippy)
library(reactable)
library(tidyr)

# load data
source("prep/library.R")
source("prep/functions.R")
source("prep/dataframes.R")

state_for_report <- "Oklahoma"
```


```{r}
################################################################################

# Highlight Section at Top

################################################################################

# Create title for prison sentencing information
prison_sentencing_system_info <- HTML("Prison Sentencing<br>System&#9432;")

# Get prison sentencing information for state
prison_sentencing_system_name <- subset(robinainfo, state == state_for_report)$prison_sentencing_system

# Reformat text depending on prison sentencing information
prison_sentencing_system <- 
  case_when(
    prison_sentencing_system_name == "Extremely High Indeterminacy" ~ HTML("Extremely High<br>Indeterminacy"),
    prison_sentencing_system_name == "High Indeterminacy"           ~ HTML("High<br>Indeterminacy"),
    prison_sentencing_system_name == "Moderate Indeterminacy"       ~ HTML("Moderate<br>Indeterminacy"),
    prison_sentencing_system_name == "Low Indeterminacy"            ~ HTML("Low<br>Indeterminacy"),
    prison_sentencing_system_name == "Extremely Low Indeterminacy"  ~ HTML("Extremely Low<br>Indeterminacy"))

# Get prison sentencing information's definition
prison_sentencing_system_definition <- subset(robinadefinitions, term == prison_sentencing_system_name)$definition 

# Assemble all prison sentencing information for the state
prison_sentencing_system_tooltip <- 
  HTML(paste0("To compare the degrees of indeterminacy, or unpredictability
   of time served, of prison-sentencing systems of different jurisdictions,
   we use a qualitative ranking framework based on the Robina Institute's
   Prison-release Discretion And Prison Population Size Report.<br><br><b>", 
   prison_sentencing_system_name, "</b>: ",
   prison_sentencing_system_definition, "."))

# Make the information accessible
prison_sentencing_system_info <- paste0('<span aria-label="', prison_sentencing_system_tooltip, '" tabindex="0" role="button">', prison_sentencing_system_info, '</span>')

# Get number if people currently eligible and eligible in the future for parole
if (state_for_report %in% unique(parole_eligibility_table_2020$state)) {
  state_data <- parole_eligibility_table_2020 %>% filter(state == state_for_report)
} else {
  state_data <- "No Data"
}
num_people_current   <- sum(state_data$current_count, na.rm = TRUE)
num_people_1_5_years <- sum(state_data$future_1_5_years_count, na.rm = TRUE)









################################################################################

# Prison Population Tab

################################################################################

# Stacked bar chart showing distribution of prison population by admission type
if (state_for_report %in% names(all_stackedbar_admtype_2020)) {
  state_stackedbar_admtype_2020 <-
    all_stackedbar_admtype_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "Distribution of the Prison Population by Admission Type")
} else {
  state_stackedbar_admtype_2020 <- "No Data"
}


###################
# Who is in Prison?
###################

# Stacked bar chart showing distribution of people by race
if (state_for_report %in% names(all_stackedbar_prison_race_2020)) {
  state_stackedbar_prison_race_2020 <-
    all_stackedbar_prison_race_2020[[state_for_report]] %>%
    hc_size(#width = 600, 
            height = 400) %>%
    hc_title(text = "Race and Ethincity")
} else {
  state_stackedbar_prison_race_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by gender
if (state_for_report %in% names(all_stackedbar_prison_gender_2020)) {
  state_stackedbar_prison_gender_2020 <-
    all_stackedbar_prison_gender_2020[[state_for_report]] %>%
    hc_size(#width = 600, 
            height = 400) %>%
    hc_title(text = "Gender")
} else {
  state_stackedbar_prison_gender_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by ageyrend
if (state_for_report %in% names(all_stackedbar_prison_ageyrend_2020)) {
  state_stackedbar_prison_ageyrend_2020 <-
    all_stackedbar_prison_ageyrend_2020[[state_for_report]] %>%
    hc_size(#width = 600, 
            height = 400) %>%
    hc_title(text = "Age")
} else {
  state_stackedbar_prison_ageyrend_2020 <- "No Data"
}


###################
# Prison Population Trends
###################

# line chart showing the change in prison pop 
# and the change in releases to parole over time
if (state_for_report %in% names(all_line_pop_released_to_parole)) {
  state_line_pop_released_to_parole <- 
    all_line_pop_released_to_parole[[state_for_report]] %>% 
    hc_size(#width = 600, 
            height = 400) %>%
    hc_title(text = "Trends in Prison Population, Releases to Parole, and Parole Eligibility")
} else {
  state_line_pop_released_to_parole <- "No Data"
}


###################
# Sentence Length
###################

# grouped barchart showing the proportion of the prison population by sentlgth and admtype
if (state_for_report %in% names(all_groupedbar_prison_sentlgth_2020)) {
  state_groupedbar_prison_sentlgth_2020 <- 
    all_groupedbar_prison_sentlgth_2020[[state_for_report]] %>% 
    hc_size(#width = 700, 
            height = 700) %>%
    hc_title(text = "Option 1: Prison Population by Sentence Length and Admission Type")
} else {
  state_groupedbar_prison_sentlgth_2020 <- "No Data"
}

# barchart showing the proportion of the prison population by sentlgth
if (state_for_report %in% names(all_bar_prison_sentlgth_2020)) {
  state_bar_prison_sentlgth_2020 <- 
    all_bar_prison_sentlgth_2020[[state_for_report]] %>% 
    hc_size(#width = 700, 
            height = 400) %>%
    hc_title(text = "Option 2: Prison Population by Sentence Length")
} else {
  state_bar_prison_sentlgth_2020 <- "No Data"
}

###################
# Criminal Offenses
###################

# grouped barchart showing the proportion of the prison population by sentlgth and admtype
if (state_for_report %in% names(all_groupedbar_prison_fbi_index_2020)) {
  state_groupedbar_prison_fbi_index_2020 <- 
    all_groupedbar_prison_fbi_index_2020[[state_for_report]] %>% 
    hc_size(#width = 700, 
      height = 700) %>%
    hc_title(text = "Option 1: Prison Population by Criminal Offense and Admission Type")
} else {
  state_groupedbar_prison_fbi_index_2020 <- "No Data"
}

# barchart showing the proportion of the prison population by sentlgth
if (state_for_report %in% names(all_bar_prison_fbi_index_2020)) {
  state_bar_prison_fbi_index_2020 <- 
    all_bar_prison_fbi_index_2020[[state_for_report]] %>% 
    hc_size(#width = 700, 
            height = 400) %>%
    hc_title(text = "Option 2: Prison Population by Criminal Offense")
} else {
  state_bar_prison_fbi_index_2020 <- "No Data"
}



################################################################################

# Parole Eligibility Tab

################################################################################

###################
# Parole-Eligible Prison Population
###################

# Stacked bar chart showing the  proportion of parole eligibility types
if (state_for_report %in% names(all_stackedbar_pe_type_2020)) {
  state_stackedbar_pe_type_2020 <- 
    all_stackedbar_pe_type_2020[[state_for_report]] %>% 
    hc_size(height = 270) %>%
    hc_title(text = "Prison Population by Parole Eligibility Status")
} else {
  state_stackedbar_pe_type_2020 <- "No Data"
}

# Save sentence about parole eligible population
if (state_for_report %in% names(all_sentence_parole_elgibility_population)) {
  state_sentence_parole_elgibility_population <- 
    all_sentence_parole_elgibility_population[[state_for_report]]
} else {
  state_sentence_parole_elgibility_population <- ""
}


###################
# How is Parole Eligibility Determined?
###################

# Get parole eligibility information by state
parole_eligibility_criteria <- subset(robinaparoleeligibility, 
                                      state == state_for_report)$general_rules_of_release_eligibility


###################
# People who are Eligible for Parole
###################

##########
# Race
##########

# Save sentence about PE by race and by state
if (state_for_report %in% names(all_sentence_parole_elgibility_race)) {
  state_sentence_parole_elgibility_race <- 
    all_sentence_parole_elgibility_race[[state_for_report]]
} else {
  state_sentence_parole_elgibility_race <- ""
}

# Save bar chart of people eligible for parole but still in prison by race and state
if (state_for_report %in% names(all_bar_parole_elgibility_race)) {
  state_bar_parole_elgibility_race <- 
    all_bar_parole_elgibility_race[[state_for_report]] %>% 
    hc_size(#width = 600, 
            height = 350) %>% 
    hc_colors(c(yellow)) %>% 
    hc_title(text = "Parole-Eligible Prison Population by Race") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_bar_parole_elgibility_race <- "No Data"
}

##########
# Gender
##########

# Save sentence about PE by gender and by state
if (state_for_report %in% names(all_sentence_parole_elgibility_gender)) {
  state_sentence_parole_elgibility_gender <- 
    all_sentence_parole_elgibility_gender[[state_for_report]]
} else {
  state_sentence_parole_elgibility_gender <- ""
}

# Save bar chart of people eligible for parole but still in prison by gender and state
if (state_for_report %in% names(all_bar_parole_elgibility_gender)) {
  state_bar_parole_elgibility_gender <- 
    all_bar_parole_elgibility_gender[[state_for_report]] %>% 
    hc_size(#width = 600, 
      height = 350) %>% 
    hc_colors(c(purple)) %>% 
    hc_title(text = "Parole-Eligible Prison Population by Gender") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_bar_parole_elgibility_gender <- "No Data"
}

##########
# Age
##########

# Save sentence about PE by ageyrend and by state
if (state_for_report %in% names(all_sentence_parole_elgibility_ageyrend)) {
  state_sentence_parole_elgibility_ageyrend <- 
    all_sentence_parole_elgibility_ageyrend[[state_for_report]]
} else {
  state_sentence_parole_elgibility_ageyrend <- ""
}

# Save bar chart of people eligible for parole but still in prison by ageyrend and state
if (state_for_report %in% names(all_bar_parole_elgibility_ageyrend)) {
  state_bar_parole_elgibility_ageyrend <- 
    all_bar_parole_elgibility_ageyrend[[state_for_report]] %>% 
    hc_size(#width = 600, 
      height = 350) %>% 
    hc_colors(c(purple)) %>% 
    hc_title(text = "Parole-Eligible Prison Population by Age") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_bar_parole_elgibility_ageyrend <- "No Data"
}




###################
# Criminal Offenses for the Parole-Eligible Prison Population
###################

# Save pie chart of most serious sentenced offense by state - new crime
if (state_for_report %in% names(all_pie_parole_elgibility_fbi_index_new_crime)) {
  state_pie_parole_elgibility_fbi_index_new_crime <- 
    all_pie_parole_elgibility_fbi_index_new_crime[[state_for_report]] %>% 
    hc_size(#width = 600, 
            height = 300) %>% 
    hc_title(text = "Option 1: Parole-Eligible Prison Population by Criminal Offense") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_pie_parole_elgibility_fbi_index_new_crime <- "No Data"
}

# Save bar chart of most serious sentenced offense by state - new crime
if (state_for_report %in% names(all_bar_parole_elgibility_fbi_index_new_crime)) {
  state_bar_parole_elgibility_fbi_index_new_crime <- 
    all_bar_parole_elgibility_fbi_index_new_crime[[state_for_report]] %>% 
    hc_size(#width = 600, 
            height = 500) %>% 
    hc_title(text = "Option 2: Parole-Eligible Prison Population by Criminal Offense") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_bar_parole_elgibility_fbi_index_new_crime <- "No Data"
}

# Save sentence about serious offense by state
if (state_for_report %in% names(all_sentence_parole_elgibility_fbi_index)) {
  state_sentence_parole_elgibility_fbi_index <- 
    all_sentence_parole_elgibility_fbi_index[[state_for_report]]
} else {
  state_sentence_parole_elgibility_fbi_index <- ""
}








################################################################################

# Releases Tab

################################################################################





################################################################################

# Disparities Tab

################################################################################








```

::: narrow-section
```{r}
########################################

# Highlights Section

########################################

htmltools::div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); justify-items: center;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_people_1_5_years, digits = 0)),
  div(style ="display: block; 
              color: #ff6400; 
              font-size: 1.25em; 
              font-family: Graphik; 
              font-weight: bold; 
              text-align: center;",
      prison_sentencing_system),
  div(id = "bold-text", HTML("Currently Eligible<br>for Parole")),
  div(id = "bold-text", HTML("Eligible for Parole<br>in the next 1-5 years")),
  div(id = "bold-text", tippy(prison_sentencing_system_info,
                            placement = "bottom",
                            flip = TRUE,
                            theme = "light",
                            tooltip = prison_sentencing_system_tooltip))
)
```
:::
<br><br>














:::: {.panel-tabset .nav-pills}

## Prison Population
<center>
```{r}
########################################

# Prison Population by Admission Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "New Court Commits vs. Parole Revocations"), br(),
    div(id = "regular-text", "While the public often envisions a prison population primarily made up of individuals serving long sentences for serious crimes, the reality is far more nuanced. In particular, the distinction between new court commits—those incarcerated for a new crime—and parole revocations—those returned to prison for violating the conditions of their parole—is a critical one. Understanding the proportion of these two groups within the prison population can provide valuable insights into the effectiveness of the state's criminal justice policies, including sentencing and parole practices."), br(),
    div(id = "visualization", state_stackedbar_admtype_2020), br()
    )
```

```{r}
########################################

# Prison Population Trends

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Prison Population Trends"), br(),
    div(id = "visualization", state_line_pop_released_to_parole), br()
    )
```

```{r}
########################################

# Who's in Prison?

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Who's in Prison?"), br(),
    div(id = "regular-text", "The following visualizations offer an overview of the demographic composition of the prison population in 2020 by race and ethnicity, gender, and age."), br(),
    )
```
::: {.panel-tabset}
### Race and Ethinicity
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_prison_race_2020)
    )
```

### Gender
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_prison_gender_2020)
    )
```

### Age
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_prison_ageyrend_2020)
    )
```
:::

```{r}
div(id = "body-section", br(),
    div(id = "big-section-title", "Sentence Length"), br(),
    div(id = "visualization", state_groupedbar_prison_sentlgth_2020), 
    div(id = "visualization", state_bar_prison_sentlgth_2020), br()
    )
```

```{r}
div(id = "body-section", br(),
    div(id = "big-section-title", "Criminal Offense"), br(),
    div(id = "visualization", state_groupedbar_prison_fbi_index_2020), 
    div(id = "visualization", state_bar_prison_fbi_index_2020), br()
    )
```

</center>
<br>













## Parole Eligibility

<center>
```{r}

########################################

# How is Parole Eligibility Determined?

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "How is Parole Eligibility Determined?"), br(),
    div(id = "regular-text",       parole_eligibility_criteria),            br(), br()
)

########################################

# Parole-Eligible Prison Population

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Parole-Eligible Prison Population"),           br(),
    div(id = "regular-text",       state_sentence_parole_elgibility_population),  br(), 
    div(id = "bar-chart",          state_stackedbar_pe_type_2020),                br()
)

########################################

# People who are Eligible for Parole

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "People who are Eligible for Parole"), br()
    )
```
::: {.panel-tabset}
### Race and Ethnicity
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  state_sentence_parole_elgibility_race), br(),
    div(id = "visualization", state_bar_parole_elgibility_race)
    )
```

### Gender
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  state_sentence_parole_elgibility_gender), br(),
    div(id = "visualization", state_bar_parole_elgibility_gender)
    )
```

### Age
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  state_sentence_parole_elgibility_ageyrend), br(),
    div(id = "visualization", state_bar_parole_elgibility_ageyrend)
    )
```

:::

```{r}
########################################

# Sentence Lengths

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Sentence Lengths"), br(),
    div(id = "regular-text", "TBD"), br()
    )
```

```{r}
########################################

# Criminal Offenses

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Criminal Offenses"), br(),
    div(id = "regular-text", state_sentence_parole_elgibility_fbi_index), br(),
    div(id = "visualization", state_pie_parole_elgibility_fbi_index_new_crime), br(),
    div(id = "visualization", state_bar_parole_elgibility_fbi_index_new_crime), br()
    )
```
</center>
<br>














## Releases
<center>
```{r}
########################################

# Release Trends

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Release Trends"), br(),
    div(id = "regular-text", "(1) # and % of parole-eligible people released each year, (2) out of all releases: % of people released each year at parole eligibility year, % 1-5 years after parole eligibility, % of people released more than 5 years after parole eligibility"), br()
    )
```

```{r}
########################################

# Demographics

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Demographics"), br(),
    div(id = "regular-text", "Gender, race, age"), br()
    )
```

```{r}
########################################

# Type of Release

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Type of Release"), br(),
    div(id = "regular-text", "(1) Unconditional vs. Conditional release, 2020, (2) Conditional vs. unconditional release, 2020"), br()
    )
```

```{r}
########################################

# Release Timing by Offense Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Release Timing by Offense Type"), br()
    )
```

```{r}
########################################

# Time Served by Offense Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Time Served by Offense Type"), br()
    )
```
</center>
<br>



















## Disparities
<center>
```{r}
########################################

# Gender

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Gender"), br(),
    div(id = "regular-text", "(1) bar charts including community, prison pop, parole eligible pop, paroled at first opportunity pop (2) RRIs for each of those populations, (3) release date compared to first eligibility date (e.g., 50% released first year, 22% year after first eligibility, 15% 2 years after eligible, etc.)"), br(),
    )
```

```{r}
########################################

# Race and Ethnicity

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Race and Ethnicity"), br(),
    div(id = "regular-text", "(1) bar charts including community, prison pop, parole eligible pop, paroled at first opportunity pop (2) RRIs for each of those populations, (3) release date compared to first eligibility date (e.g., 50% released first year, 22% year after first eligibility, 15% 2 years after eligible, etc.)"), br(),
    )
```

```{r}
########################################

# Predicted Probabilities

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Predicted Probabilities"), br(),
    div(id = "regular-text", "Predicted probabilities: Controlling for offense and sentence length, predicted probabilities of release before/at eligibility date.  Something like “After controlling for differences in ages, offenses, and sentence length, 66% of White people, 40% of Black people, and 55% of Hispanic people are released before/on parole eligibility date. Similarly, 80% of women and 74% of men were released before/on parole eligibility date"), br(),
    )
```

</center>
<br>

::::






::: citation-section

Citations:   

Alper, Mariel E., et al. "Profiles in Parole Release and Revocation: Examining the Legal Framework in the United States." *Robina Institute of Criminal Law and Criminal Justice*, 13 May 2022, [robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states](robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states).     

Carson, Ann E. "Prisoners in 2020 – Statistical Tables." Bureau of Justice Statistics, 1 Dec. 2020, [bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables](bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables).    

United States. Bureau of Justice Statistics. Annual Parole Survey, 2018. Inter-university Consortium for Political and Social Research [distributor], 2021-10-28. [https://doi.org/10.3886/ICPSR38058.v1](https://doi.org/10.3886/ICPSR38058.v1). 
  
United States. Bureau of Justice Statistics. National Corrections Reporting Program, 1991-2020: Selected Variables. Inter-university Consortium for Political and Social Research [distributor], 2022-11-28. [https://doi.org/10.3886/ICPSR38492.v1](https://doi.org/10.3886/ICPSR38492.v1).     
:::

<br><br>

---
title: "`r state_for_report`"
---
