---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/config.R")

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/carl_state_notes.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_info_by_state.rds"))

# Parole Eligibility Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_pe_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_ped_offense_group.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_ped_fbi_index.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_parole_eligibility_sentlgth.rds"))

# Population Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_line_population_by_year.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_population_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_population_fbi_index.rds"))

# Releases Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_releases.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_line_releases_by_year.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_pie_release_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_agerlse.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_parole_eligibility_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_offense_los.rds"))


# Disparities Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_parole_release_disparities.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_rri_chart.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_rri.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_black.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_hispanic.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_white.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_other.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_los_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_los_race_offense.rds"))

state_for_report <- "Idaho"

source("R/pull_state_findings.R")
```

<div id="state-body-section">

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("Pct. of People in Prison<br>Past Their Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board Members"))
)
br()
```

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Year

`r paste(state_sentence_parole_eligibility_population, collapse = ", ")`

```{r}
div(id = "visualization", state_stackedbar_pe_type)
```

### Demographics

The demographics of people in prison past their parole eligibility year reveals notable proportions among non-Hispanic Black and White people. Gender distribution indicates a predominance of males over females. Age-wise, the majority of individuals fell within the 25-34 years and 35-44 years brackets.

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_parole_eligibility_race),
  div(id = "small-visualization", state_waffle_parole_eligibility_sex),
  div(id = "small-visualization", state_waffle_parole_eligibility_ageyrend),
)
```

### Criminal Offense

```{r}
# section_text <- HTML("")
# 
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 10px; margin-bottom: 0px;",
#   section_text,
#   div(id = "visualization", state_bubble_ped_offense_group)
# )
```

In 2020, 61% of people in prison past their parole eligibility date were in prison for violent offenses.

```{r}
div(id = "highlight-visualization", style = "width: 200px; height: 200px; overflow: hidden;", state_bubble_ped_offense_group)
br()
```

The breakdown of criminal offenses of people in prison past their parole eligibility year reveals a varied landscape, with the majority of people incarcerated for aggravated or simple assault (26%) and property crimes (19%). Other offense categories make up smaller proportions of this population.

```{r}
div(id = "visualization", state_bar_ped_fbi_index)
```


### Sentence Length

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br>
<div class="divider"></div>

## Missing Data

This section will have information about missing parole eligibility data for the state once new NCRP data is released.

<br>



# Prison Population

## Prison Population Trends

`r paste(state_sentence_population, collapse = ", ")`

```{r}
div(id = "visualization", state_line_population_by_year)
```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners Series</a>

<br>
<div class="divider"></div>

## Demographics

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_population_race),
  div(id = "small-visualization", state_waffle_population_sex),
  div(id = "small-visualization", state_waffle_population_ageyrend)
)
```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners 2022 - Race and Ethnicity</a>

<br>
<div class="divider"></div>

## Criminal Offenses

```{r}
div(id = "visualization", state_bar_population_fbi_index)
```


<br>
<div class="divider"></div>

## Sentence Lengths

```{r}
div(id = "visualization", state_bar_population_sentlgth)
```

<br>








# Prison Releases

## Prison Release Trends

`r paste(state_sentence_releases, collapse = ", ")`

```{r}
div(id = "visualization", state_line_releases_by_year)
```
<br>

The proportions of parole-eligible individuals released versus those not released have shown fluctuating trends from 2010 to 2020. In 2020, 40% of parole-eligible individuals were released, while 60% remained in prison, reflecting ongoing challenges and changes in parole policies and practices.  

```{r}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. In 2020, 45% of releases are conditional, requiring ongoing supervision and compliance, while 55% are unconditional, allowing individuals to reintegrate without additional constraints.

```{r}
div(id = "visualization", state_pie_release_type)
```


<br>
<div class="divider"></div>

## Demographics of People Released

The demographics of people released from prison in 2020 reveals notable proportions among non-Hispanic Black and White people. Gender distribution indicates a predominance of males over females. Age-wise, the majority of individuals fall within the 25-34 years and 35-44 years brackets. These findings provide insights into the populations transitioning back into the community.

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_releases_race),
  div(id = "small-visualization", state_waffle_releases_sex),
  div(id = "small-visualization", state_waffle_releases_agerlse),
)
```

<br>
<div class="divider"></div>

## Time Served by Offense

Between 2010 and 2020, shifts in average time served by individuals for different offense types have been observed. While some categories, like murder and non-negligent manslaughter, have seen a substantial increase in time served, others, like drug-related offenses, have remained relatively stable. 

```{r}
div(id = "visualization", state_lollipop_offense_los)
```

<br>












# Disparities

## Incarceration Rates

`r paste(state_sentence_rri, collapse = ", ")`

```{r}
div(id = "visualization", state_hc_waffle_rri_black)
div(id = "visualization", state_hc_waffle_rri_white)
div(id = "visualization", state_hc_waffle_rri_hispanic)
div(id = "visualization", state_hc_waffle_rri_other)
```


<br>
<div class="divider"></div>

## Years Spent in Prison After Parole Eligibility

```{r, results='asis'}
df1 <- all_parole_release_disparities |> filter(state == state_for_report) |> 
  select(time_between_ped_release, race) |> 
  group_by(race) |> 
  summarise(total_years = sum(time_between_ped_release, na.rm = TRUE))

message <- paste("In 2020, people in prison past their parole eligibility year collectively spent a total of",
                 comma(sum(df1$total_years)), " years behind bars instead of supervised in the community on parole. This extended incarceration varied across different racial and ethnic groups.")

cat(message)
```

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
  div(style = "color: #00bdb2; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Black, non-Hispanic"]) ),
  div(style = "color: #3f8bca; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "White, non-Hispanic"]) ),
  div(style = "color: #de663e; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Hispanic, any race"]) ),
  div(style = "color: #f4b811; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Other race(s), non-Hispanic"]) ),
  div(style = "color: #00bdb2; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
  div(style = "color: #3f8bca; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
  div(style = "color: #de663e; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
  div(style = "color: #f4b811; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
  div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Black People")),
  div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("White People")),
  div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Hispanic People")),
  div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Other Race(s), Non-Hispanic People"))
)
br()
```

```{r}
div(id = "visualization", state_scatter_race_ped_release)
```

<br>
<div class="divider"></div>

## Timing of Release After Parole Eligibility

The timing of release for individuals after their parole eligibility year varied by offense type and demographic group, with many remaining in prison for several years beyond their eligibility date. 

```{r}
div(id = "visualization", state_bubble_race_ped_release)
```

<br>
<div class="divider"></div>

## Time Served

```{r}
section_text <- "Hispanic individuals faced the longest average length of stay in prison in 2020. Similarly, Black, non-Hispanic individuals also experienced notably longer prison stays compared to their White, non-Hispanic counterparts, who had the shortest average length of stay."
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 10px;",
  section_text,
  div(id = "visualization", state_lollipop_los_race)
)
br()
```

```{r}
div(id = "visualization", state_scatter_los_race_offense)
```


<br>
<br>




:::

</div>

---
title: "`r state_for_report`"
---
