---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/config.R")

# Load Prepared Data --------------------------------------------------
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/carl_state_notes.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_info_by_state.rds"))

# Parole Eligibility Tab --------------------------------------------------
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_pe_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_demographics.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_fbi_index.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_ped_offense_group.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_ped_fbi_index.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_parole_eligibility_sentlgth.rds"))

# Population Tab --------------------------------------------------
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_line_population_by_year.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_population_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_population_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_population_fbi_index.rds"))

# Releases Tab --------------------------------------------------
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_releases.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_line_releases_by_year.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_pie_release_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_releases_demographics.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_releases_agerlse.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_pe_proportion_released.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_parole_eligibility_release.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_los_offense.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_offense_los.rds"))


# Disparities Tab --------------------------------------------------
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_parole_release_disparities.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_rri_chart.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_rri.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_black.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_hispanic.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_white.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_hc_waffle_rri_other.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_los_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_los_race.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_los_race_offense.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_los_race_offense.rds"))

state_for_report <- "Georgia"

source("R/pull_state_findings.R")
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("Prison Population Incarcerated Past Their Parole Eligibility Year")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board<br>Members"))
)
br()
```

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Year

`r paste(state_sentence_parole_eligibility_population, collapse = ", ")`

```{r state_stackedbar_pe_type}
div(id = "visualization", state_stackedbar_pe_type)
```

### Demographics

`r paste(state_sentence_parole_eligibility_demographics, collapse = ", ")`

```{r state_waffle_parole_eligibility_demographics}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_parole_eligibility_race),
  div(id = "small-visualization", state_waffle_parole_eligibility_sex),
  div(id = "small-visualization", state_waffle_parole_eligibility_ageyrend),
)
```

### Criminal Offenses

```{r state_sentence_parole_eligibility_fbi_index}
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: 3fr 1fr; grid-row-gap: 0px; grid-column-gap: 20px;",
#   state_sentence_parole_eligibility_fbi_index,
#   div(id = "highlight-visualization", style = "width: 200px; height: 200px; overflow: hidden; margin-right: 0; margin-left: 0px;", state_bubble_ped_offense_group)
# )
# br()
```

`r paste(state_sentence_parole_eligibility_fbi_index, collapse = ", ")`

```{r state_bar_ped_fbi_index}
# div(id = "visualization", state_bubble_ped_offense_group)
div(id = "visualization", state_bar_ped_fbi_index)
```


### Original Sentence Lengths

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r state_bar_parole_eligibility_sentlgth}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br>
<div class="divider"></div>

## Missing Data

This section will contain information related to missingness in the NCRP data. Keep?
<br>
<br>
<br>










# Prison Population

## Prison Population Trends

`r paste(state_sentence_population, collapse = ", ")`

```{r state_line_population_by_year}
div(id = "visualization", state_line_population_by_year)
```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners Series</a>

<br>
<div class="divider"></div>

## Demographics

```{r state_waffle_population_demographics}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_population_race),
  div(id = "small-visualization", state_waffle_population_sex),
  div(id = "small-visualization", state_waffle_population_ageyrend)
)
```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners 2022 - Race and Ethnicity</a>

<br>
<div class="divider"></div>

## Criminal Offenses

```{r state_bar_population_fbi_index}
div(id = "visualization", state_bar_population_fbi_index)
```


<br>
<div class="divider"></div>

## Sentence Lengths

```{r state_bar_population_sentlgth}
div(id = "visualization", state_bar_population_sentlgth)
```

<br>








# Prison Releases

## Prison Release Trends

`r paste(state_sentence_releases, collapse = ", ")`

```{r state_line_releases_by_year}
div(id = "visualization", state_line_releases_by_year)
```
<br>

`r paste(state_sentence_pe_proportion_released, collapse = ", ")`

```{r state_stackedbar_parole_eligibility_release}
div(id = "visualization", state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

Conditional release involves an individual’s release under specific conditions and supervision, whereas unconditional release means the individual is released without further obligations or restrictions. 

```{r state_pie_release_type}
div(id = "visualization", state_pie_release_type)
```


<br>
<div class="divider"></div>

## Demographics of People Released

`r paste(state_sentence_releases_demographics, collapse = ", ")`

```{r state_waffle_releases_demographics}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_releases_race),
  div(id = "small-visualization", state_waffle_releases_sex),
  div(id = "small-visualization", state_waffle_releases_agerlse),
)
```

<br>
<div class="divider"></div>

## Time Served by Offense

`r paste(state_sentence_los_offense, collapse = ", ")`

```{r state_lollipop_offense_los}
div(id = "visualization", state_lollipop_offense_los)
```

<br>












# Disparities

## Incarceration Rates

`r paste(state_sentence_rri, collapse = ", ")`

```{r state_hc_waffle_rri_demographics}
div(id = "visualization", state_hc_waffle_rri_black)
div(id = "visualization", state_hc_waffle_rri_white)
div(id = "visualization", state_hc_waffle_rri_hispanic)
div(id = "visualization", state_hc_waffle_rri_other)
```


<br>
<div class="divider"></div>

## Years Spent in Prison After Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r state_parole_release_disparities, results='asis'}
df1 <- all_parole_release_disparities |>
  filter(state == state_for_report) |> 
  filter(!is.na(race)) |> 
  select(time_between_ped_release, race) |> 
  group_by(race) |> 
  summarise(total_years = sum(time_between_ped_release, na.rm = TRUE))

# Check if all total_years are NA
if (!all(is.na(df1$total_years))) {
  
  message <- HTML(paste0("In ", select_year, ", people in prison past their parole eligibility year collectively spent a total of ",
                   comma(sum(df1$total_years, na.rm = TRUE)), " years behind bars instead of supervised in the community on parole. This extended incarceration varied across different racial and ethnic groups."))
  
  cat(message)
  
  br()
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Black, non-Hispanic"]) ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "White, non-Hispanic"]) ),
    div(style = "color: #d97d68; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Hispanic, any race"]) ),
    div(style = "color: #decf64; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Other race(s), non-Hispanic"]) ),
    div(style = "color: #49a7a1; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #55b4e5; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #d97d68; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #decf64; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Black People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("White People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Hispanic People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Other Race(s), Non-Hispanic People"))
  )
  
  cat(as.character("<br><br>"))
  cat(as.character(html_output))
  br()
} else {
  cat("")
}
```

```{r state_scatter_race_ped_release}
div(id = "visualization", state_scatter_race_ped_release)
```

### Gender

In progress. Same type of graph as race and ethnicity tab.

:::

<br>
<div class="divider"></div>




## Timing of Release After Parole Eligibility

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

The timing of release for individuals after their parole eligibility year varied by offense type and demographic group, with many remaining in prison for several years beyond their eligibility date. 

```{r state_bubble_race_ped_release}
div(id = "visualization", state_bubble_race_ped_release)
```

### Gender

In progress. Same type of graph as race and ethnicity tab.

:::

<br>
<div class="divider"></div>




## Time Served

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r}
# section_text <- "X individuals faced the longest average time served in prison in 2020. Similarly, Y, individuals also experienced notably longer prison stays compared to their Z counterparts, who had the shortest average time served."
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   section_text,
#   div(id = "visualization", state_lollipop_los_race)
# )
# br()
```

`r paste(state_sentence_los_race, collapse = ", ")`

```{r state_lollipop_los_race}
# div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_los_race)
div(id = "visualization", style = "margin: 0 auto; text-align: center;", state_lollipop_los_race)
```

`r paste(state_sentence_los_race_offense, collapse = ", ")`

```{r state_scatter_los_race_offense}
div(id = "visualization", state_scatter_los_race_offense)
```

### Gender 

In progress. Same type of graph as race and ethnicity tab.

:::

<br>
<br>




:::

</div>

---
title: "`r state_for_report`"
---
