---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

# This will change depending on the state
state_for_report <- "Oklahoma"

# Load config file and pregenerated sentences and visualizations
source("R/config.R")
source("R/pull_state_findings.R")

# Function to display data and sentences
# Display no data if data is missing
fnc_display_sentence_chart <- function(sentence, chart, chart_type = "visualization", final_text = no_data_text) {
  # Check if both sentence and chart are missing
  no_data <- is.null(sentence) || is.null(chart) 
  
  # Display the sentence and chart if available
  if (!no_data) {
    cat(sentence)
    div(id = chart_type, chart)
  } else {
    # Show "no data" message only once if both are missing
    final_text
  }
}
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(proj_ppey, digits = 0)),
  div(id = "highlight-text", paste0(round(proj_pcnt_ppey, 0), "%")),
  div(id = "highlight-text", formattable::comma(parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("People In Prison Past<br>Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Prison Population Incarcerated<br>Past Parole Eligibility")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Parole Board<br>Members<sup style='font-size: 0.5em;'>1</sup>"))
)
```

<br>
<div id="asterisk-section"> `r additional_asterisks_combined` </div><br>
</div>

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility Is Determined

Parole eligibility refers to the point in a person's prison sentence when they become eligible for consideration for release under parole. Often, there is a minimum term a person must serve before becoming parole eligible. Parole is a conditional release, allowing the person to serve the remainder of their sentence in the community under supervision, rather than in prison.  

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Parole Eligibility

```{r results = 'asis', state_sentence_pe_type}
fnc_display_sentence_chart(state_sentence_pe_type, state_pie_pe_type); br()
fnc_display_sentence_chart(state_sentence_pop_pe_by_year, state_line_pop_pe_by_year)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_pe_race}
fnc_display_sentence_chart(state_sentence_pe_race, state_bar_pe_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_pe_sex}
fnc_display_sentence_chart(state_sentence_pe_sex, state_bar_pe_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_pe_ageyrend}
fnc_display_sentence_chart(state_sentence_pe_ageyrend, state_bar_pe_ageyrend,
                           chart_type = "small-visualization")
```

:::

<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_pe_fbi_index}
fnc_display_sentence_chart(state_sentence_pe_fbi_index, state_bar_pe_fbi_index)
```

<br>
<div class="divider"></div>

## Sentence Lengths

```{r results = 'asis', state_sentence_pe_sentlgth}
fnc_display_sentence_chart(state_sentence_pe_sentlgth, state_bar_pe_sentlgth)
```

<br>
<div class="divider"></div>

## Estimation Methodology

`r paste(state_imputation_notes, collapse = ", ")`

<br>
<div class="divider"></div>

## Endnotes

`r paste(state_citation, collapse = ", ")`

Data analysis was conducted using information from the National Corrections Reporting Program and the National Prisoner Statistics Program. See “National Corrections Reporting Program (NCRP),” Bureau of Justice Statistics, accessed September 25, 2024, [https://bjs.ojp.gov/data-collection/national-corrections-reporting-program-ncrp](https://bjs.ojp.gov/data-collection/national-corrections-reporting-program-ncrp); “Prisoners - Statistical Tables,” Bureau of Justice Statistics, accessed September 25, 2024, [https://bjs.ojp.gov/library/publications/list?series_filter=Prisoners](https://bjs.ojp.gov/library/publications/list?series_filter=Prisoners).






```{r}
#-----------------------------------------------------------------------------------------------------------------------------------------#
```

# Prison Population

## Prison Population Trends

```{r results = 'asis', state_sentence_population}
fnc_display_sentence_chart(state_sentence_population, state_line_population_by_year)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_population_race}
fnc_display_sentence_chart(state_sentence_population_race, state_bar_population_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_population_sex}
fnc_display_sentence_chart(state_sentence_population_sex, state_bar_population_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_population_ageyrend}
fnc_display_sentence_chart(state_sentence_population_ageyrend, state_bar_population_ageyrend,
                           chart_type = "small-visualization")
```

:::

<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_population_fbi_index}
fnc_display_sentence_chart(state_sentence_population_fbi_index, state_bar_population_fbi_index)
```

<br>
<div class="divider"></div>

## Sentence Lengths

```{r results = 'asis', state_sentence_population_sentlgth}
fnc_display_sentence_chart(state_sentence_population_sentlgth, state_bar_population_sentlgth)
```













```{r}
#-----------------------------------------------------------------------------------------------------------------------------------------#
```


# Prison Releases

## Prison Release Trends

```{r results = 'asis', state_sentence_releases}
fnc_display_sentence_chart(state_sentence_releases, state_line_releases_by_year); br()
fnc_display_sentence_chart(state_sentence_pe_proportion_released, state_stackedbar_pe_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

```{r results = 'asis', state_sentence_release_type }
fnc_display_sentence_chart(state_sentence_release_type, state_pie_release_type)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_releases_race}
fnc_display_sentence_chart(state_sentence_releases_race, state_bar_releases_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_releases_sex}
fnc_display_sentence_chart(state_sentence_releases_sex, state_bar_releases_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_releases_agerlse}
fnc_display_sentence_chart(state_sentence_releases_agerlse, state_bar_releases_agerlse,
                           chart_type = "small-visualization")
```

:::
  
<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_releases_fbi_index}
fnc_display_sentence_chart(state_sentence_releases_fbi_index, state_bar_releases_fbi_index)
```












```{r}
#-----------------------------------------------------------------------------------------------------------------------------------------#
```


# Disparities

## Measuring Disparities

Disparities can be measured in many different ways by comparing the outcomes of different groups. This page details disparities in three types of metrics by race, ethnicity, and sex:   

(1\) the proportion of the incarcerated population who are past their parole eligibility date,<br>
(2\) the average number of years individuals spent behind bars past their parole eligibility date,<br>
(3\) the average time individuals served in prison, and<br>
(4\) the average time individuals served in prison by offense type.<br>

<br>
<div class="divider"></div>

## Disparities in People Incarcerated Past Parole Eligibility

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r}
no_infographic_black <- !file.exists(pe_rri_infographic_black)
no_data_black <- !state_for_report %in% all_pe_rri_data$state
rri_details_black <- all_pe_rri_data |> 
  filter(state == state_for_report & race == "Black, non-Hispanic") |> 
  group_by(state, race) |> 
  summarise(
    above_1 = rri > 1,
    equal_1 = rri == 1,
    below_1 = rri < 1
  ) |> 
  mutate(rri_exists = ifelse(equal_1 == FALSE, TRUE, FALSE))
no_rri_details_black <- nrow(rri_details_black) == 0

no_infographic_hispanic <- !file.exists(pe_rri_infographic_hispanic)
no_data_hispanic <- !state_for_report %in% all_pe_rri_data$state
rri_details_hispanic <- all_pe_rri_data |> 
  filter(state == state_for_report & race == "Hispanic, any race") |> 
  group_by(state, race) |> 
  summarise(
    above_1 = rri > 1,
    equal_1 = rri == 1,
    below_1 = rri < 1
  ) |> 
  mutate(rri_exists = ifelse(equal_1 == FALSE, TRUE, FALSE))
no_rri_details_hispanic <- nrow(rri_details_hispanic) == 0

# Define whether to use "Other race(s), non-Hispanic" based on state
use_other_race <- state_for_report %in% states_use_other_race_eth$state

# Check for the "Other race(s), non-Hispanic" infographic and data
no_infographic_other <- !file.exists(pe_rri_infographic_other)
no_data_other <- !state_for_report %in% all_pe_rri_data$state
rri_details_other <- all_pe_rri_data |> 
  filter(state == state_for_report & race == "Other race(s), non-Hispanic") |> 
  group_by(state, race) |> 
  summarise(
    above_1 = rri > 1,
    equal_1 = rri == 1,
    below_1 = rri < 1
  ) |> 
  mutate(rri_exists = ifelse(equal_1 == FALSE, TRUE, FALSE))
no_rri_details_other <- nrow(rri_details_other) == 0
```

```{r results = 'asis'}
if (no_infographic_black == FALSE) {
  HTML("<br>")
}
if (no_infographic_black == TRUE & no_data_black == TRUE) {
  # Display "No Data" message
  no_data_needed_black_individuals <- TRUE
} else if (no_infographic_black == TRUE & no_data_black == FALSE & no_rri_details_black == FALSE) {
  # Display "No RRI" message
  HTML(paste0("<br><span style='font-weight:bold; color:#49a7a1;'>Black, non-Hispanic people</span> were incarcerated in state prison past their parole eligibility at the same rate as <span style='font-weight:bold; color:#d97d68;'>White people</span>, when accounting for prison population sizes in ", state_for_report, ".<br><br>"))
} else if (no_infographic_black == FALSE & no_data_black == FALSE) {
  # Render the image using knitr::include_graphics()
  knitr::include_graphics(pe_rri_infographic_black)
}

if (no_infographic_black == FALSE & no_data_black == FALSE) {
  # Render the HTML content separately
  HTML(state_sentence_pe_rri_black, "<br><br>")
}

if (no_infographic_hispanic == TRUE & no_data_hispanic == TRUE) {
  # Display "No Data" message
  no_data_needed_hispanic_individuals <- TRUE
} else if (no_infographic_hispanic == TRUE & no_data_hispanic == FALSE & no_rri_details_hispanic == FALSE) {
  # Display "No RRI" message
  HTML(paste0("<span style='font-weight:bold; color:#55b4e5;'>Hispanic people</span> were incarcerated in state prison past their parole eligibility at the same rate as <span style='font-weight:bold; color:#d97d68;'>White people</span>, when accounting for prison population sizes in ", state_for_report, ".<br>"))
} else if (no_infographic_hispanic == FALSE & no_data_hispanic == FALSE) {
  # Render the image using knitr::include_graphics()
  knitr::include_graphics(pe_rri_infographic_hispanic)
}
if (no_infographic_hispanic == FALSE & no_data_hispanic == FALSE) {
  # Render the HTML content separately
  HTML(state_sentence_pe_rri_hispanic, "<br>")
}

# Only process "Other race(s), non-Hispanic" if the state is in the specified list
if (use_other_race) {
  if (no_infographic_other == TRUE & no_data_other == TRUE) {
    no_data_needed_other_individuals <- TRUE
  } else if (no_infographic_other == TRUE & no_data_other == FALSE & no_rri_details_other == FALSE) {
    HTML(paste0("<br><span style='font-weight:bold; color:#938ebf;'>Non-Hispanic people of other race(s)</span> were incarcerated in state prison past their parole eligibility at the same rate as <span style='font-weight:bold; color:#d97d68;'>White people</span>, when accounting for prison population sizes in ", state_for_report, ".<br>"))
  } else if (no_infographic_other == FALSE & no_data_other == FALSE) {
    knitr::include_graphics(pe_rri_infographic_other)
  }

  if (no_infographic_other == FALSE & no_data_other == FALSE) {
    HTML(state_sentence_pe_rri_other, "<br>")
  }
}

# Display "No Data" message if all three groups (Black, Hispanic, Other) have no data
if ((no_infographic_black == TRUE & no_data_black == TRUE) &
    (no_infographic_hispanic == TRUE & no_data_hispanic == TRUE) &
    (!use_other_race | (use_other_race & no_infographic_other == TRUE & no_data_other == TRUE))) {
  no_data_text
} else {
  cat(as.character(div(id = "rri_sourcenotes-text", br(), ncrp_csg_source_year)))
}
```

### Sex

```{r results = 'asis'}
no_infographic_male <- !file.exists(pe_rri_infographic_male)
no_data_male <- !state_for_report %in% all_pe_rri_data_male$state
rri_details_male <- all_pe_rri_data_male |> 
  filter(state == state_for_report & sex == "Male") |> 
  group_by(state, sex) |> 
  summarise(
    above_1 = rri > 1,
    equal_1 = rri == 1,
    below_1 = rri < 1
  ) |> 
  mutate(rri_exists = ifelse(equal_1 == FALSE, TRUE, FALSE))
no_rri_details_male <- nrow(rri_details_male) == 0

if (no_infographic_male == FALSE) {
  HTML("<br>")
}
if (no_infographic_male == TRUE & no_data_male == TRUE) {
  # Display "No Data" message
  no_data_needed_male_individuals <- TRUE
} else if (no_infographic_male == TRUE & no_data_male == FALSE & no_rri_details_male == FALSE) {
  # Display "No RRI" message
  HTML(paste0(
    "<br><span style='font-weight:bold; color:#49a7a1;'>Men</span> were incarcerated in state prison past their parole eligibility at the same rate as 
     <span style='font-weight:bold; color:#938ebf;'>Women</span>, when accounting for prison population sizes in ", 
     state_for_report, ".<br>"
  ))
} else if (no_infographic_male == FALSE & no_data_male == FALSE) {
  # Render the image using knitr::include_graphics()
  knitr::include_graphics(pe_rri_infographic_male)
} 

if (no_infographic_male == FALSE & no_data_male == FALSE) {
  HTML(state_sentence_pe_rri_male, "<br>")
}

# Display "No Data" message if no data
if (no_infographic_male == TRUE & no_data_male == TRUE){
  no_data_text
} else {
  cat(as.character(div(id = "rri_sourcenotes-text", br(), ncrp_csg_source_year)))
}
```

:::



<br>
<div class="divider"></div>

## Years in Prison Past Parole Eligibility

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r results = 'asis'}
no_data <- is.null(state_sentence_avg_past_pe_race)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_race, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_avg_past_pe_race)
}

if (!is.null(state_scatter_avg_past_pe_race_offense)) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_race_offense, "<br><br>"))
}

if (!is.null(state_scatter_avg_past_pe_race_offense)) {
  div(id = "visualization", state_scatter_avg_past_pe_race_offense)
}

if (no_data) {
  no_data_text
}
```


### Sex

```{r results = 'asis'}
no_data <- is.null(state_sentence_avg_past_pe_sex)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_sex, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_avg_past_pe_sex)
}

if (!is.null(state_scatter_avg_past_pe_sex_offense)) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_sex_offense, "<br><br>"))
}

if (!is.null(state_scatter_avg_past_pe_sex_offense)) {
  div(id = "visualization", state_scatter_avg_past_pe_sex_offense)
}

if (no_data) {
  no_data_text
}
```

:::

<br>
<div class="divider"></div>

## Time Served

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r results='asis'}
no_data <- is.null(state_sentence_los_race)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_los_race, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_los_race)
}

if (!is.null(state_scatter_los_race_offense)) {
  HTML(paste0("<br>", state_sentence_los_race_offense, "<br><br>"))
}

if (!is.null(state_scatter_los_race_offense)) {
  div(id = "visualization", state_scatter_los_race_offense)
}

if (no_data) {
  no_data_text
}
```



### Sex

```{r results='asis'}
no_data <- is.null(state_sentence_los_sex)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_los_sex, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_los_sex)
}

if (!is.null(state_scatter_los_sex_offense)) {
  HTML(paste0("<br>", state_sentence_los_sex_offense, "<br><br>"))
}

if (!is.null(state_scatter_los_sex_offense)) {
  div(id = "visualization", state_scatter_los_sex_offense)
}

if (no_data) {
  no_data_text
}
```

:::



```{r}
#-----------------------------------------------------------------------------------------------------------------------------------------#
```

<br>
<div class="divider"></div>

## Notes on Disparities

**Understanding Disparities in Parole Outcomes**

In examining disparities in parole release rates, this analysis uses the Relative Rate Index (RRI) as a metric to assess differences in incarceration rates past parole eligibility across race, ethnicity, and sex.

The RRI is calculated by dividing the rate for Black or Hispanic individuals by the rate for White individuals, with values above 1.0 indicating that a group experiences the outcome at a higher rate relative to White individuals, and values below 1.0 indicating a lower rate. This allows for a comparative analysis of which groups are more likely to remain incarcerated past their parole eligibility.

**Metrics Displayed in Visualizations**

Our visualizations present three main metrics to highlight disparities across race, ethnicity, and sex:

- **RRI of Parole Release:** This metric compares the proportion of individuals past parole eligibility within the total prison population by race, ethnicity, and sex. It provides a snapshot of who remains in prison despite being eligible for parole.
- **Average Time Past Parole Eligibility:** For those individuals who remain incarcerated past their parole eligibility date, we show the average additional time spent in prison. This is broken down by race, ethnicity, and sex, highlighting the length of delay in parole release among different groups.
- **Average Total Time Served:** We also display the average total time served in prison by race, ethnicity, and sex, providing insight into sentencing and incarceration length disparities.

**Data Considerations**
Our analysis focuses on Black, Hispanic, and White categories to align with data limitations from the National Corrections Reporting Program (NCRP), which categorizes individuals as White, Black, Hispanic, or Other. The term “Hispanic” is used here to align with NCRP terminology.

Data collection and reporting practices vary significantly across states:

- Race and Ethnicity Identification: Some states collect race and ethnicity data based on intake officers’ assessments, while others allow individuals to self-identify. This may result in variability and influence RRI calculations.
- Combined vs. Separate Categories: Some states combine race and ethnicity into a single category, while others separate them. This inconsistency can impact the accuracy and comparability of data, sometimes leading to an underreporting of Hispanic individuals and mixed-race individuals or an overrepresentation of non-Hispanic White individuals.
- Including 'Other' Category in Analysis: In states where populations categorized as "Other race(s), non-Hispanic" are substantial, we include these groups in the analysis. 

**Small Sample Sizes and Interpretation**

Interpret findings with small numbers cautiously, as they may introduce variability and yield less reliable estimates. We exclude individuals categorized as "Other race(s), non-Hispanic" when the sample size is too small to make meaningful comparisons. According to the NCRP, the “Other race(s)” category may include American Indian or Alaska Native, Asian, Native Hawaiian or Other Pacific Islander, and individuals identifying as more than one race. Small sample sizes can produce unreliable or potentially misleading results in disparities analysis.
:::

</div>






---
title: "`r state_for_report`"
---
