---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)


source("R/config.R")

# Load prepared data
load(file = paste0(config$sp_data_path, "/data/analysis/app/map_percent.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_eligibility_table.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/carl_state_notes.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/parole_info_by_state.rds"))

# Parole Eligibility Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_population.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_stackedbar_pe_type.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_race.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_sex.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_waffle_parole_eligibility_ageyrend.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_ped_fbi_index.rds"))

load(file = paste0(config$sp_data_path, "/data/analysis/app/all_sentence_parole_eligibility_sentlgth.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bar_parole_eligibility_sentlgth.rds"))

# Releases Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_lollipop_offense_los.rds"))

# Disparities Tab
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_scatter_race_ped_release.rds"))
load(file = paste0(config$sp_data_path, "/data/analysis/app/all_bubble_race_ped_release.rds"))


state_for_report <- "Georgia"

source("R/pull_state_findings.R")
```

<div id="state-body-section">

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", paste0(round(num_people_current_perc*100, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", HTML("Pct. of People in Prison<br>Past Their Parole Eligibility Date")),
  div(id = "bold-text", style = "text-align: center;", HTML("In Prison Past Their<br>Parole Eligibility Date")),
  div(id = "bold-text", style = "text-align: center;", HTML("Parole Board Members"))
)
br()
```

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility is Determined

`r paste(parole_eligibility_criteria, collapse = ", ")`
<div class="divider"></div>

## In Prison Past Their Parole Eligibility Date

`r paste(state_sentence_parole_eligibility_population, collapse = ", ")`

```{r}
div(id = "visualization", state_stackedbar_pe_type)
```

### Demographics

The charts below provide further details on the demographics of this population, illustrating the percentage of people in prison past their parole eligibility date by race and ethnicity, gender, and current age.  

```{r}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-gap: 20px;",
  div(id = "small-visualization", state_waffle_parole_eligibility_race),
  div(id = "small-visualization", state_waffle_parole_eligibility_sex),
  div(id = "small-visualization", state_waffle_parole_eligibility_ageyrend),
)
```


### Criminal Offense

Here, the charts present data on the types of criminal offenses committed by people in prison past their parole eligibility date.  

```{r}
div(id = "visualization", state_bubble_ped_fbi_index)
```


### Sentence Length

`r paste(state_sentence_parole_eligibility_sentlgth, collapse = ", ")`

```{r}
div(id = "visualization", state_bar_parole_eligibility_sentlgth)
```

<br>
<div class="divider"></div>

## Missing Data

This section will have information about missing parole eligibility data for the state once new NCRP data is released.

<br>



# Prison Population

The size and composition of the prison population have significant implications for state resources and public safety. 

Sentence about change in year-end population from 2010 to 2020.  
Trend graph

<br>
<div class="divider"></div>

## Demographics

An examination of the demographic composition of Georgia's prison population, including age, gender, race, and other relevant factors.  

<br>
<div class="divider"></div>

## Criminal Offenses

An overview of the types of criminal offenses that lead to incarceration, providing insight into the most common crimes and their prevalence.  

<br>
<div class="divider"></div>

## Sentence Lengths

Analysis of the distribution of sentence lengths within the prison population.  










# Prison Releases

Understanding trends in prison releases is crucial for assessing the effectiveness of parole and reentry programs.  

Sentence about change in releases from 2010 to 2020.
Trend graph

Proportion of parole-eligible populations released each year.

## Prison Release Types

Prison releases that were conditional and unconditional.

<br>
<div class="divider"></div>

## Demographics

A demographic analysis of individuals released from prison, highlighting trends and patterns in reentry.  

<br>
<div class="divider"></div>

## Timing of Release

An exploration of the timing of prison releases, including how long individuals serve before being released.  

Timing of Release Relative to Parole Eligibility Year by Release Type (Unconditional and Conditional)

<br>
<div class="divider"></div>

## Time Served by Offense Type

Detailed analysis of the average time served by individuals based on their offense type.  

```{r}
div(id = "visualization", state_lollipop_offense_los)
```














# Disparities

Examining disparities within the prison system is essential for addressing issues of equity and justice. This section focuses on disparities in incarceration and release practices.  

In prison. X People are X times more likely to be incarcerated than White people.

## In Prison Past Their Parole Eligibility Date

<br>
<div class="divider"></div>

## Years Spent in Prison After Parole Eligibility

ADD TOTAL HERE

This graph depicts the years individuals spend in prison after becoming eligible for parole by race and ethnicity. It highlights disparities in parole release timings among various racial and ethnic groups.  

```{r}
div(id = "visualization", state_scatter_race_ped_release)
```

<br>
<div class="divider"></div>

## Timing of Release by Offense Type

This chart illustrates the proportion of parole-eligible people released from prison in their first, second, or third year or more after becoming eligible for parole. The data is categorized by offense type and racial and ethnic group, revealing that release patterns can vary for the same offense across these groups.

```{r}
div(id = "visualization", state_bubble_race_ped_release)
```

<br>
<div class="divider"></div>

## Time Served 

<br>
<div class="divider"></div>






:::

</div>

---
title: "`r state_for_report`"
---
