---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

# This will change depending on the state
state_for_report <- "Georgia"

# Load config file and pregenerated sentences and visualizations
source("R/config.R")
source("R/pull_state_findings.R")

# Function to display data and sentences
# Display no data if data is missing
fnc_display_sentence_chart <- function(sentence, chart, chart_type = "visualization", final_text = no_data_text) {
  # Check if both sentence and chart are missing
  no_data <- is.null(sentence) || is.null(chart) 
  
  # Display the sentence and chart if available
  if (!no_data) {
    cat(sentence)
    div(id = chart_type, chart)
  } else {
    # Show "no data" message only once if both are missing
    final_text
  }
}
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(proj_ppey, digits = 0)),
  div(id = "highlight-text", paste0(round(proj_pcnt_ppey, 0), "%")),
  div(id = "highlight-text", formattable::comma(parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("People In Prison Past<br>Parole Eligibility*")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Prison Population Incarcerated<br>Past Parole Eligibility")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Parole Board<br>Members<sup style='font-size: 0.5em;'>1</sup>"))
)
```

<br>
<div id="asterisk-section"> `r additional_asterisks_combined` </div><br>

::: {.panel-tabset .nav-pills}
 
# Parole Eligibility

```{r}
render_image <- JS("
  function(){
    this.renderer.image('https://avparoleproject.netlify.app/img/csgjc-logo.png', 10, this.chartHeight - 50, 140.1, 30)
    .add();
  }")

x <- c(rnorm(10000), rnorm(1000, 4, 0.5))

hc <- hchart(x, name = "data", color = "#17b8b6") 

hc |> 
  hc_add_dependency(name = "modules/exporting.js") |>
  hc_add_dependency(name = "modules/offline-exporting.js") |>
  hc_exporting(
      enabled = TRUE,
      buttons = list(contextButton = list(menuItems = list("downloadPNG"))),
      chartOptions = list(
        chart = list(
          style = list(fontFamily = "Helvetica"),
          events = list(
            load = render_image
            )
          )
        )
      ) |> 
  hc_chart(
    style = list(fontWeight = 400, fontFamily = "Times New Roman")
  )
```


## How Parole Eligibility Is Determined

Parole eligibility refers to the point in a person's prison sentence when they become eligible for consideration for release under parole. Often, there is a minimum term a person must serve before becoming parole eligible. Parole is a conditional release, allowing the person to serve the remainder of their sentence in the community under supervision, rather than in prison.  

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Parole Eligibility

```{r results = 'asis', state_sentence_pe_type}
fnc_display_sentence_chart(state_sentence_pe_type, state_pie_pe_type); br()
fnc_display_sentence_chart(state_sentence_pop_pe_by_year, state_line_pop_pe_by_year)
```

<br>
<div class="divider"></div>

:::

</div>
 





---
title: "`r state_for_report`"
---
