---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
    anchor-sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

# This will change depending on the state
state_for_report <- "South Dakota"

# Load config file and pregenerated sentences and visualizations
source("R/config.R")
source("R/pull_state_findings.R")

# Function to display data and sentences
# Display no data if data is missing
fnc_display_sentence_chart <- function(sentence, chart, chart_type = "visualization", final_text = no_data_text) {
  # Check if both sentence and chart are missing
  no_data <- is.null(sentence) || is.null(chart) 
  
  # Display the sentence and chart if available
  if (!no_data) {
    cat(sentence)
    div(id = chart_type, chart)
  } else {
    # Show "no data" message only once if both are missing
    final_text
  }
}
```

<div id="state-body-section">

```{r highlight}
div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", paste0(round(num_people_current_perc, 0), "%")),
  div(id = "highlight-text", formattable::comma(num_parole_board_mem, digits = 0)),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("People In Prison Past<br>Parole Eligibility*<br><span style='font-weight: normal; font-size: 0.8em'>(2023 Projection TBD)</span>")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Prison Population Incarcerated<br>Past Parole Eligibility<br><span style='font-weight: normal; font-size: 0.8em'>(2023 Projection TBD)</span>")),
  div(id = "bold-text", style = "text-align: center;", 
      HTML("Parole Board<br>Members"))
)
```

<br>
<div id="asterisk-section"> `r footnote_text` </div><br>
</div>

::: {.panel-tabset .nav-pills}

# Parole Eligibility

## How Parole Eligibility Is Determined

Parole eligibility refers to the point in a person's prison sentence when they become eligible for consideration for release under parole. Parole is a conditional release, allowing the person to serve the remainder of their sentence in the community under supervision, rather than in prison.  

`r paste(parole_eligibility_criteria, collapse = ", ")`

<br>
<div class="divider"></div>

## In Prison Past Parole Eligibility

```{r results = 'asis', state_sentence_pe_type}
fnc_display_sentence_chart(state_sentence_pe_type, state_pie_pe_type); br()
fnc_display_sentence_chart(state_sentence_pop_pe_by_year, state_bar_pop_pe_by_year)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_parole_eligibility_race}
fnc_display_sentence_chart(state_sentence_parole_eligibility_race, state_bar_parole_eligibility_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_parole_eligibility_sex}
fnc_display_sentence_chart(state_sentence_parole_eligibility_sex, state_bar_parole_eligibility_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_parole_eligibility_ageyrend}
fnc_display_sentence_chart(state_sentence_parole_eligibility_ageyrend, state_bar_parole_eligibility_ageyrend,
                           chart_type = "small-visualization")
```

:::

<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_parole_eligibility_fbi_index}
fnc_display_sentence_chart(state_sentence_parole_eligibility_fbi_index, state_bar_ped_fbi_index)
```

<br>
<div class="divider"></div>

## Sentence Lengths

```{r results = 'asis', state_sentence_parole_eligibility_sentlgth}
fnc_display_sentence_chart(state_sentence_parole_eligibility_sentlgth, state_bar_parole_eligibility_sentlgth)
```

<br>
<div class="divider"></div>

## Estimation Methodology

`r paste(state_imputation_notes, collapse = ", ")`

<br>
<div class="divider"></div>

## Endnotes

`r paste(state_citation, collapse = ", ")`

Data analysis was conducted using information from the National Corrections Reporting Program and the National Prisoner Statistics Program. See “National Corrections Reporting Program (NCRP),” Bureau of Justice Statistics, accessed September 25, 2024, [https://bjs.ojp.gov/data-collection/national-corrections-reporting-program-ncrp](https://bjs.ojp.gov/data-collection/national-corrections-reporting-program-ncrp); “Prisoners - Statistical Tables,” Bureau of Justice Statistics, accessed September 25, 2024, [https://bjs.ojp.gov/library/publications/list?series_filter=Prisoners](https://bjs.ojp.gov/library/publications/list?series_filter=Prisoners).






```{r}
####################################################################################################################
```

# Prison Population

## Prison Population Trends

```{r results = 'asis', state_sentence_population}
fnc_display_sentence_chart(state_sentence_population, state_line_population_by_year)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_population_race}
fnc_display_sentence_chart(state_sentence_population_race, state_bar_population_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_population_sex}
fnc_display_sentence_chart(state_sentence_population_sex, state_bar_population_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_population_ageyrend}
fnc_display_sentence_chart(state_sentence_population_ageyrend, state_bar_population_ageyrend,
                           chart_type = "small-visualization")
```

:::

<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_population_fbi_index}
fnc_display_sentence_chart(state_sentence_population_fbi_index, state_bar_population_fbi_index)
```

<br>
<div class="divider"></div>

## Sentence Lengths

```{r results = 'asis', state_sentence_population_sentlgth}
fnc_display_sentence_chart(state_sentence_population_sentlgth, state_bar_population_sentlgth)
```













```{r}
####################################################################################################################
```

# Prison Releases

## Prison Release Trends

```{r results = 'asis', state_sentence_releases}
fnc_display_sentence_chart(state_sentence_releases, state_line_releases_by_year); br()
fnc_display_sentence_chart(state_sentence_pe_proportion_released, state_stackedbar_parole_eligibility_release)
```

<br>
<div class="divider"></div>

## Conditional Releases

```{r results = 'asis', state_sentence_release_type }
fnc_display_sentence_chart(state_sentence_release_type, state_pie_release_type)
```

<br>
<div class="divider"></div>

## Demographics

::: {.panel-tabset .custom-tabs}

### Race and Ethnicity

```{r results = 'asis', state_sentence_releases_race}
fnc_display_sentence_chart(state_sentence_releases_race, state_bar_releases_race,
                           chart_type = "small-visualization")
```

### Sex

```{r results = 'asis', state_sentence_releases_sex}
fnc_display_sentence_chart(state_sentence_releases_sex, state_bar_releases_sex,
                           chart_type = "small-visualization")
```

### Age

```{r results = 'asis', state_sentence_releases_agerlse}
fnc_display_sentence_chart(state_sentence_releases_agerlse, state_bar_releases_agerlse,
                           chart_type = "small-visualization")
```

:::
  
<br>
<div class="divider"></div>

## Offense Types

```{r results = 'asis', state_sentence_releases_fbi_index}
fnc_display_sentence_chart(state_sentence_releases_fbi_index, state_bar_releases_fbi_index)
```












```{r}
####################################################################################################################
```

# Disparities

## Disparities in Parole Release

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r results = 'asis'}
# Check if both Black and Hispanic data are missing
no_data_black <- !file.exists(pe_rri_infographic_black) 
no_data_hispanic <- !file.exists(pe_rri_infographic_hispanic) 

if (file.exists(pe_rri_infographic_black)) {
  HTML("<br>")
}
if (file.exists(pe_rri_infographic_black)) {
  knitr::include_graphics(pe_rri_infographic_black)
}
if (!no_data_black) {
  HTML(state_sentence_pe_rri_black, "<br><br>")
}

if (file.exists(pe_rri_infographic_hispanic)) {
  knitr::include_graphics(pe_rri_infographic_hispanic)
}
if (!no_data_hispanic) {
  HTML(state_sentence_pe_rri_hispanic)
}

if (no_data_black && no_data_hispanic) {
  no_data_text
}
```

### Sex

```{r results = 'asis'}
no_data_male <- !file.exists(pe_rri_infographic_male) 

if (file.exists(pe_rri_infographic_male)) {
  HTML("<br>")
}
if (file.exists(pe_rri_infographic_male)) {
  knitr::include_graphics(pe_rri_infographic_male)
}
if (!no_data_male) {
  HTML(paste0(state_sentence_pe_rri_male, "<br>"))
}

if (no_data_male) {
  no_data_text
}
```

:::



<br>
<div class="divider"></div>

## Years in Prison Past Parole Eligibility

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r results = 'asis'}
df1 <- avg_current_pe_race |>
  filter(state == state_for_report)

no_data <- nrow(df1) == 0

if (!no_data) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_race, "<br>"))
}

if (!no_data) {
    html_output <- div(
    id = "grid-container",
    # style = "display: grid; grid-template-columns: repeat(3, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
        style = "display: grid; grid-template-columns: repeat(3, auto); justify-content: center; align-items: center; text-align: center; column-gap: 50px;",  # Reduced column gap to bring them closer
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;",
        paste0(round(df1$avg_years_to_estimated_pey[df1$race == "Black, non-Hispanic"], 1), " Years")
        ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;",
        paste0(round(df1$avg_years_to_estimated_pey[df1$race == "Hispanic, any race"], 1), " Years")
        ),
    div(style = "color: #d97d68; text-align: center; font-size: 24px; font-weight: bold;",
        paste0(round(df1$avg_years_to_estimated_pey[df1$race == "White, non-Hispanic"], 1), " Years")
        ),

    div(style = "color: #49a7a1; text-align: center; font-size: 16px; font-weight: bold;", HTML("Black")),
    div(style = "color: #55b4e5; text-align: center; font-size: 16px; font-weight: bold;", HTML("Hispanic")),
    div(style = "color: #d97d68; text-align: center; font-size: 16px; font-weight: bold;", HTML("White"))
  )
  cat(as.character(br()))
  cat(as.character(div(id = "bold-text", "Average Years in Prison Past Parole Eligibility", br())))
  cat(as.character(html_output))
}

if (!is.null(state_scatter_avg_past_pe_race_offense)) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_race_offense, "<br><br>"))
}

if (!is.null(state_scatter_avg_past_pe_race_offense)) {
  div(id = "visualization", state_scatter_avg_past_pe_race_offense)
}

if (no_data) {
  no_data_text
}
```


### Sex

```{r results = 'asis'}
df1 <- avg_current_pe_sex |>
  filter(state == state_for_report)

no_data <- nrow(df1) == 0

if (!no_data) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_sex, "<br>"))
}

if (!no_data) {
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(2, auto); justify-content: center; align-items: center; text-align: center; column-gap: 50px;",  # Reduced column gap to bring them closer
    # First column: Years for Male, Second column: Years for Female
    div(style = "color: #49a7a1; font-size: 24px; font-weight: bold; margin: 0;", 
        paste0(round(df1$avg_years_to_estimated_pey[df1$sex == "Male"], 1), " Years")
    ),
    div(style = "color: #55b4e5; font-size: 24px; font-weight: bold; margin: 0;", 
        paste0(round(df1$avg_years_to_estimated_pey[df1$sex == "Female"], 1), " Years")
    ),
    # First column: Male label, Second column: Female label
    div(style = "color: #49a7a1; font-size: 16px; font-weight: bold; margin: 0;", HTML("Male")),
    div(style = "color: #55b4e5; font-size: 16px; font-weight: bold; margin: 0;", HTML("Female"))
  )
  cat(as.character(br()))
  cat(as.character(div(id = "bold-text", "Average Years in Prison Past Parole Eligibility", br())))
  cat(as.character(html_output))
}

if (!is.null(state_scatter_avg_past_pe_sex_offense)) {
  HTML(paste0("<br>", state_sentence_avg_past_pe_sex_offense, "<br><br>"))
}

if (!is.null(state_scatter_avg_past_pe_sex_offense)) {
  div(id = "visualization", state_scatter_avg_past_pe_sex_offense)
}

if (no_data) {
  no_data_text
}
```

:::

```{r results = 'asis'}
# KEEP THIS CODE FOR NOW IN CASE TEAM WANTS IT BACK IN
# <br>
# <div class="divider"></div>
# 
# ## Disparities in Incarceration
# 
# ::: {.panel-tabset .custom-tabs-disp}
# 
# ### Race and Ethnicity
# 
# # Check if both Black and Hispanic data are missing
# no_data_black <- !file.exists(rri_infographic_black) 
# no_data_hispanic <- !file.exists(rri_infographic_hispanic) 
# 
# if (file.exists(rri_infographic_black)) {
#   HTML("<br>")
# }
# if (file.exists(rri_infographic_black)) {
#   knitr::include_graphics(rri_infographic_black)
# }
# if (!no_data_black) {
#   HTML(paste0(state_sentence_rri_black, "<br><br>"))
# }
# if (file.exists(rri_infographic_hispanic)) {
#   knitr::include_graphics(rri_infographic_hispanic)
# }
# if (!no_data_hispanic) {
#   HTML(paste0(state_sentence_rri_hispanic, "<br>"))
# }
# 
# if (no_data_black && no_data_hispanic) {
#   no_data_text
# }
# ```
# 
# ### Sex
# ```{r results = 'asis'}
# # no_data_male <- !file.exists(rri_infographic_male) 
# # 
# # if (file.exists(rri_infographic_male)) {
# #   HTML("<br>")
# # }
# # if (file.exists(rri_infographic_male)) {
# #   knitr::include_graphics(rri_infographic_male)
# # }
# # if (!no_data_male) {
# #   HTML(paste0(state_sentence_rri_male, "<br>"))
# # }
# # 
# # if (no_data_male) {
# #   no_data_text
# # }
# ```
# 
# :::
```

<br>
<div class="divider"></div>

## Time Served

::: {.panel-tabset .custom-tabs-disp}

### Race and Ethnicity

```{r results='asis'}
no_data <- is.null(state_sentence_los_race)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_los_race, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_los_race)
}

if (!is.null(state_scatter_los_race_offense)) {
  HTML(paste0("<br>", state_sentence_los_race_offense, "<br><br>"))
}

if (!is.null(state_scatter_los_race_offense)) {
  div(id = "visualization", state_scatter_los_race_offense)
}

if (no_data) {
  no_data_text
}
```



### Sex

```{r results='asis'}
no_data <- is.null(state_sentence_los_sex)

if (!no_data) {
  HTML(paste0("<br>", state_sentence_los_sex, "<br><br>"))
}

if (!no_data) {
  div(id = "small-visualization", state_lollipop_los_sex)
}

if (!is.null(state_scatter_los_sex_offense)) {
  HTML(paste0("<br>", state_sentence_los_sex_offense, "<br><br>"))
}

if (!is.null(state_scatter_los_sex_offense)) {
  div(id = "visualization", state_scatter_los_sex_offense)
}

if (no_data) {
  no_data_text
}
```

:::





<br>
<div class="divider"></div>

## Notes on Disparities

**Understanding Disparities in Parole Outcomes**

In examining disparities in parole release rates, this analysis uses the Relative Rate Index (RRI) as a metric to assess differences in criminal justice outcomes—including incarceration rates, parole eligibility, and time served—across race, ethnicity, and sex.

The RRI is calculated by dividing the rate for Black or Hispanic individuals by the rate for White individuals, with values above 1.0 indicating that a group experiences the outcome at a higher rate relative to White individuals, and values below 1.0 indicating a lower rate. This allows for a comparative analysis of which groups are more likely to remain incarcerated past their parole eligibility.

**Metrics Displayed in Visualizations**

Our visualizations present three main metrics to highlight disparities across race, ethnicity, and sex:

- **RRI of Parole Release:** This metric compares the proportion of individuals past parole eligibility within the total prison population by race, ethnicity, and sex. It provides a snapshot of who remains in prison despite being eligible for parole.
- **Average Time Past Parole Eligibility:** For those individuals who remain incarcerated past their parole eligibility date, we show the average additional time spent in prison. This is broken down by race, ethnicity, and sex, highlighting the length of delay in parole release among different groups.
- **Average Total Time Served:** We also display the average total time served in prison by race, ethnicity, and sex, providing insight into sentencing and incarceration length disparities.

**Data Considerations**
Our analysis focuses on Black, Hispanic, and White categories to align with data limitations from the National Corrections Reporting Program (NCRP), which categorizes individuals as White, Black, Hispanic, or Other. The term “Hispanic” is used here to align with NCRP terminology.

Data collection and reporting practices vary significantly across states:

- Race and Ethnicity Identification: Some states collect race and ethnicity data based on intake officers’ assessments, while others allow individuals to self-identify. This may result in variability and influence RRI calculations.
- Combined vs. Separate Categories: Some states combine race and ethnicity into a single category, while others separate them. This inconsistency can impact the accuracy and comparability of data, sometimes leading to an underreporting of Hispanic individuals and mixed-race individuals or an overrepresentation of non-Hispanic White individuals.

**Small Sample Sizes and Interpretation**

Interpret findings with small numbers cautiously, as they may introduce variability and yield less reliable estimates. We exclude individuals categorized as "Other race(s), non-Hispanic" because the small numbers in this group limit the ability to make meaningful comparisons. According to NCRP, the “Other race(s)” category may include American Indian or Alaska Native, Asian, Native Hawaiian or Other Pacific Islander, and individuals identifying as more than one race. Small sample sizes can produce unreliable or potentially misleading results in disparities analysis.

:::





```{r}
####################################################################################################################
```

</div>







---
title: "`r state_for_report`"
---
