---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png"
)

# load packages
library(shiny)
library(highcharter)
library(dplyr)
library(htmlwidgets)
library(dplyr)
library(htmltools)
library(reactable)
library(reactablefmtr)
library(htmltools)
library(tippy)
library(reactable)
library(tidyr)

# load data
source("prep/00_library.R")
source("prep/01_functions.R")
source("prep/dataframes.R")

state_for_report <- "Kansas"
```


```{r}
# create title for prison sentencing information
prison_sentencing_system_info <- HTML("Prison Sentencing<br>System&#9432;")

# get prison sentencing information for state
prison_sentencing_system_name <- subset(robinainfo, state == state_for_report)$prison_sentencing_system

# reformat text depending on prison sentencing information
prison_sentencing_system <- 
  case_when(
    prison_sentencing_system_name == "Extremely High Indeterminacy" ~ HTML("Extremely High<br>Indeterminacy"),
    prison_sentencing_system_name == "High Indeterminacy"           ~ HTML("High<br>Indeterminacy"),
    prison_sentencing_system_name == "Moderate Indeterminacy"       ~ HTML("Moderate<br>Indeterminacy"),
    prison_sentencing_system_name == "Low Indeterminacy"            ~ HTML("Low<br>Indeterminacy"),
    prison_sentencing_system_name == "Extremely Low Indeterminacy"  ~ HTML("Extremely Low<br>Indeterminacy"))

# get prison sentencing information's definition
prison_sentencing_system_definition <- subset(robinadefinitions, term == prison_sentencing_system_name)$definition 

# assemble all prison sentencing information for the state
prison_sentencing_system_tooltip <- 
  HTML(paste0("To compare the degrees of indeterminacy, or unpredictability
   of time served, of prison-sentencing systems of different jurisdictions,
   we use a qualitative ranking framework based on the Robina Institute's
   Prison-release Discretion And Prison Population Size Report.<br><br><b>", 
   prison_sentencing_system_name, "</b>: ",
   prison_sentencing_system_definition, "."))

# make the information accessible
prison_sentencing_system_info <- paste0('<span aria-label="', prison_sentencing_system_tooltip, '" tabindex="0" role="button">', prison_sentencing_system_info, '</span>')

# get parole eligibility information
parole_eligibility_criteria <- subset(robinaparoleeligibility, state == state_for_report)$general_rules_of_release_eligibility
```


```{r}
# data prep by tab

################################################################################

# Highlights

################################################################################

if (state_for_report %in% unique(parole_eligibility_table_2020$state)) {
  state_data <- parole_eligibility_table_2020 %>% filter(state == state_for_report)
} else {
  state_data <- "No Data"
}

num_people_current   <- sum(state_data$current_count, na.rm = TRUE)
num_people_1_5_years <- sum(state_data$future_1_5_years_count, na.rm = TRUE)






################################################################################

# Prison Population

################################################################################

# line chart showing the change in prison pop 
# and the change in releases to parole over time
if (state_for_report %in% names(all_line_pop_released_to_parole)) {
  state_line_pop_released_to_parole <- 
    all_line_pop_released_to_parole[[state_for_report]] %>% 
    hc_size(width = 600, height = 400) %>%
    hc_title(text = "Trends in Prison Population, Releases to Parole, and Parole Eligibility")
} else {
  state_line_pop_released_to_parole <- "No Data"
}

# stacked bar chart showing the  proportion of parole eligibility types
if (state_for_report %in% names(all_stackedbar_pe_type_2020)) {
  state_stackedbar_pe_type_2020 <- 
    all_stackedbar_pe_type_2020[[state_for_report]] %>% 
    hc_size(height = 200) %>%
    hc_title(text = "Distribution of Parole Eligibility Status Among the Prison Population")
} else {
  state_stackedbar_pe_type_2020 <- "No Data"
}

# save sentence about parole eligible population
if (state_for_report %in% names(all_sentence_parole_elgibility_population)) {
  state_sentence_parole_elgibility_population <- 
    all_sentence_parole_elgibility_population[[state_for_report]]
} else {
  state_sentence_parole_elgibility_population <- ""
}






################################################################################

# Parole Eligibility

################################################################################

# save bar chart of people eligible for parole but still in prison by race and state
if (state_for_report %in% names(all_bar_parole_elgibility_race)) {
  state_bar_parole_elgibility_race <- 
    all_bar_parole_elgibility_race[[state_for_report]] %>% 
    hc_size(width = 600, height = 350) %>% 
    hc_title(text = "People who are Parole Eligible but Still in Prison by Race in 2020")
} else {
  state_bar_parole_elgibility_race <- "No Data"
}

# save sentence about PE by race and by state
if (state_for_report %in% names(all_sentence_parole_elgibility_race)) {
  state_sentence_parole_elgibility_race <- 
    all_sentence_parole_elgibility_race[[state_for_report]]
} else {
  state_sentence_parole_elgibility_race <- ""
}


###################
# Offenses
###################

# save sentence about parole eligiblity by adm type
if (state_for_report %in% names(all_sentence_parole_elgibility_admtype)) {
  state_sentence_parole_elgibility_admtype <- 
    all_sentence_parole_elgibility_admtype[[state_for_report]]
} else {
  state_sentence_parole_elgibility_admtype <- ""
}

# save pie chart of currently eligible
if (state_for_report %in% names(all_pie_ped_current)) {
  state_pie_ped_current <- 
    all_pie_ped_current[[state_for_report]] %>% 
    # hc_chart(width = 250, height = 350) %>% 
    hc_title(text = "Currently Eligible for Parole but in Prison due to a Parole Revocation")
} else {
  state_pie_ped_current <- ""
}

# save pie chart of future 1-5 years eligible
if (state_for_report %in% names(all_pie_ped_future_1_5)) {
  state_pie_ped_future_1_5 <- 
    all_pie_ped_future_1_5[[state_for_report]] %>% 
    # hc_chart(width = 250, height = 350) %>% 
    hc_title(text = "Eligible for Parole in 1-5 Years but in Prison due to a Parole Revocation")
} else {
  state_pie_ped_future_1_5 <- ""
}

# save pie chart of futyre 6+ years eligible
if (state_for_report %in% names(all_pie_ped_future_6)) {
  state_pie_ped_future_6 <- 
    all_pie_ped_future_6[[state_for_report]] %>% 
    hc_chart(width = 300, height = 300) %>% 
    hc_title(text = "Eligible for Parole in 6+ Years")
} else {
  state_pie_ped_future_6 <- ""
}




# save pie chart of most serious sentenced offense by state
if (state_for_report %in% names(all_pie_parole_elgibility_offense)) {
  state_pie_parole_elgibility_offense <- 
    all_pie_parole_elgibility_offense[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Most Serious Sentenced Offense for People who are Parole Eligible but Still in Prison in 2020")
} else {
  state_pie_parole_elgibility_offense <- "No Data"
}

# save pie chart of most serious sentenced offense by state - new crime
if (state_for_report %in% names(all_pie_parole_elgibility_offense_new_crime)) {
  state_pie_parole_elgibility_offense_new_crime <- 
    all_pie_parole_elgibility_offense_new_crime[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Most Serious Sentenced Offense for People who are Parole Eligible but Still in Prison in 2020")
} else {
  state_pie_parole_elgibility_offense_new_crime <- "No Data"
}

# save pie chart of most serious sentenced offense by state - parole revocation or return
if (state_for_report %in% names(all_pie_parole_elgibility_offense_parole_return)) {
  state_pie_parole_elgibility_offense_parole_return <- 
    all_pie_parole_elgibility_offense_parole_return[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Most Serious Sentenced Offense for People who are Parole Eligible but Still in Prison in 2020")
} else {
  state_pie_parole_elgibility_offense_parole_return <- "No Data"
}

# save sentence about serious offense by state
if (state_for_report %in% names(all_sentence_parole_elgibility_offense)) {
  state_sentence_parole_elgibility_offense <- 
    all_sentence_parole_elgibility_offense[[state_for_report]]
} else {
  state_sentence_parole_elgibility_offense <- ""
}





# save sentence duration by by state
if (state_for_report %in% names(all_bar_sentence_overview_2020)) {
  state_bar_sentence_overview_2020 <- 
    all_bar_sentence_overview_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for All Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_overview_2020 <- "No Data"
}

# save sentence duration by offense type (violent) and by state
if (state_for_report %in% names(all_bar_sentence_violent_2020)) {
  state_bar_sentence_violent_2020 <- 
    all_bar_sentence_violent_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for Violent Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_violent_2020 <- "No Data"
}

# save sentence duration by offense type (drugs) and by state
if (state_for_report %in% names(all_bar_sentence_drugs_2020)) {
  state_bar_sentence_drugs_2020 <- 
    all_bar_sentence_drugs_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for Drug Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_drugs_2020 <- "No Data"
}

# save sentence duration by offense type (property) and by state
if (state_for_report %in% names(all_bar_sentence_property_2020)) {
  state_bar_sentence_property_2020 <- 
    all_bar_sentence_property_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for Property Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_property_2020 <- "No Data"
}

# save sentence duration by offense type (public order) and by state
if (state_for_report %in% names(all_bar_sentence_publicorder_2020)) {
  state_bar_sentence_publicorder_2020 <- 
    all_bar_sentence_publicorder_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for Public Order Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_publicorder_2020 <- "No Data"
}

# save sentence duration by offense type (other) and by state
if (state_for_report %in% names(all_bar_sentence_other_2020)) {
  state_bar_sentence_other_2020 <- 
    all_bar_sentence_other_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Sentence Duration for Other Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_sentence_other_2020 <- "No Data"
}






################################################################################

# Releases from Prison

################################################################################

# save adm type by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_2020)) {
  state_bar_released_at_ped_2020 <- 
    all_bar_released_at_ped_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for All Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_2020 <- "No Data"
}

# save other offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_violent_2020)) {
  state_bar_released_at_ped_violent_2020 <- 
    all_bar_released_at_ped_violent_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for Violent Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_violent_2020 <- "No Data"
}

# save publicorder offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_publicorder_2020)) {
  state_bar_released_at_ped_publicorder_2020 <- 
    all_bar_released_at_ped_publicorder_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for Public Order Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_publicorder_2020 <- "No Data"
}

# save property offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_property_2020)) {
  state_bar_released_at_ped_property_2020 <- 
    all_bar_released_at_ped_property_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for Property Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_property_2020 <- "No Data"
}

# save drug offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_drugs_2020)) {
  state_bar_released_at_ped_drugs_2020 <- 
    all_bar_released_at_ped_drugs_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for Drug Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_drugs_2020 <- "No Data"
}

# save other offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_other_2020)) {
  state_bar_released_at_ped_other_2020 <- 
    all_bar_released_at_ped_other_2020[[state_for_report]] %>%
    hc_size(height = 300, width = 500) %>%
    hc_title(text = "Timing of Release for Other Offenses: Parole Revocation vs. New Crime")
} else {
  state_bar_released_at_ped_other_2020 <- "No Data"
}






################################################################################

# Disparities

################################################################################

#####
# Relative Rate Index (RRI)
#####

# graph showing RRI for Black people
if (state_for_report %in% names(all_bar_rri_sentence_length_black)) {
  state_bar_rri_sentence_length_black <- 
    all_bar_rri_sentence_length_black[[state_for_report]] %>% 
    hc_size(width = 300, height = 300) 
} else {
  state_bar_rri_sentence_length_black <- "No Data"
}

# graph showing RRI for Hispanic people
if (state_for_report %in% names(all_bar_rri_sentence_length_hispanic)) {
  state_bar_rri_sentence_length_hispanic <- 
    all_bar_rri_sentence_length_hispanic[[state_for_report]] %>% 
    hc_size(width = 300, height = 300) 
} else {
  state_bar_rri_sentence_length_hispanic <- "No Data"
}



# # graph showing the time from PED to release by race and ethnicity
# if (state_for_report %in% names(all_time_between_release_ped_2020_by_race)) {
#   state_time_between_release_ped_by_race <- 
#     all_time_between_release_ped_2020_by_race[[state_for_report]] %>% 
#     hc_title(text = "Timing of Release for People Eligible for Parole by Race and Ethnicity in 2020") %>% 
#     hc_size(height = 300, width = 500)
# } else {
#   state_time_between_release_ped_by_race <- "No Data"
# }


#####
# Predicted Probabilities
#####

# define function to create reactable tables for different predicted probabilities
fnc_create_reactable <- function(df, variable_name, variable_pp, col_name) {
  
  df$variable_name <- get(variable_name, df)
  df$variable_pp <- get(variable_pp, df)
  
  df1 <- df %>%
    filter(state == state_for_report) %>%
    select(variable_name, variable_pp) %>%
    distinct()

  reactable(df1,
            style = hc_reactable_style,
            theme = hc_reactable_theme,
            defaultColDef = colDef(format = colFormat(separators = TRUE),
                                   align = "left"),
            compact = TRUE,
            fullWidth = FALSE,
            columns = list(
              variable_name = colDef(name = col_name, minWidth = 200),
              variable_pp = colDef(name = "",
                                   minWidth = 175,
                                   align = "right",
                                   format = colFormat(percent = TRUE,
                                                      digits = 0)))
  )
}

# apply function to different variables
# race
table_pp_by_race <- 
  fnc_create_reactable(all_pp_by_variable, 
                       "race", "pp_by_race", "Race/Ethnicity")

# sex
table_pp_by_sex <- 
  fnc_create_reactable(all_pp_by_variable, 
                       "sex", "pp_by_sex", "Gender")

# offense
table_pp_by_offgeneral <- 
  fnc_create_reactable(all_pp_by_variable, 
                       "offgeneral", "pp_by_offgeneral", "Most Serious Sentenced Offense")

# sentence length
table_pp_by_sentlgth <- 
  fnc_create_reactable(all_pp_by_variable, 
                       "sentlgth", "pp_by_sentlgth", "Maximum Sentence Length")

```

::: narrow-section
```{r}
########################################

# Highlights

########################################
htmltools::div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); justify-items: center;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_people_1_5_years, digits = 0)),
  div(style ="display: block; 
              color: #ff6400; 
              font-size: 1.25em; 
              font-family: Graphik; 
              font-weight: bold; 
              text-align: center;",
      prison_sentencing_system),
  div(id = "bold-text", HTML("Currently Eligible for<br>Parole but in Prison")),
  div(id = "bold-text", HTML("Eligible for Parole in<br>the next 1-5 years")),
  div(id = "bold-text", tippy(prison_sentencing_system_info,
                              placement = "bottom",
                              flip = TRUE,
                              theme = "light",
                              tooltip = prison_sentencing_system_tooltip))
)
```
:::
<br>














:::: {.panel-tabset .nav-pills}

## Parole Eligibility

<center>
```{r}
########################################

# Parole-Eligible Prison Population

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Parole-Eligible Prison Population"),             br(),
    div(id = "regular-text",       state_sentence_parole_elgibility_population),    br(), 
    div(id = "bar-chart",          state_stackedbar_pe_type_2020),                  br(), br()
)

########################################

# How is Parole Release Eligibility Determined?

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "How is Parole Release Eligibility Determined?"), br(),
    div(id = "regular-text",       parole_eligibility_criteria),                    br(), br()
)
```

```{r}
########################################

# People who are Eligible for Parole

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "People who are Eligible for Parole"), br()
    )
```
::: {.panel-tabset}
### Race
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  state_sentence_parole_elgibility_race), br(),
    div(id = "visualization", state_bar_parole_elgibility_race),      br()
    )
```

### Gender

### Age
:::



```{r}
########################################

# Most Serious Sentenced Offense

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Most Serious Sentenced Offense"), br()
    )
```

::: {.panel-tabset}
### Overview
```{r}
div(id = "body-section", br(),
    div(id = "regular-text", state_sentence_parole_elgibility_offense), br(),
    state_pie_parole_elgibility_offense)
```

### New Crime
```{r}
div(id = "body-section",br(),
    state_pie_parole_elgibility_offense_new_crime)
```

### Parole Return
```{r}
div(id = "body-section", br(),
    state_pie_parole_elgibility_offense_parole_return)
```
:::
</center>
<br>

















## Prison Population

<center>
```{r}
########################################

# Prison Population

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Prison Population"), br(),
    div(id = "visualization",      state_line_pop_released_to_parole), br()
    )
```

```{r}
########################################

# Projected Prison Population

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Projected Prison Population"), br(),
    div(id = "visualization",     "Same as chart above but projected prison population if people who were eligible for parole were released."), br()
    )
```

</center>
<br>


















## Releases from Prison

<center>
```{r}
########################################

# Releases from Prison

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Releases from Prison"), br(),
    "TBD Visualization: Conditional vs Unconditional Releases", br(), br(),
    "TBD Visualization: Percentage/number of people who maxed out even though they were parole eligible. Mandatory release date vs actual release date, supervised vs unsupervised release.",
    br(), br(), br(), br(), 
    div(id = "big-section-title", "Timing of Release by Offense Type in 2020"), br()
    )
```

::: {.panel-tabset}
### Overview
```{r}
div(id = "body-section", state_bar_released_at_ped_2020)
```

### Violent
```{r}
div(id = "body-section", state_bar_released_at_ped_violent_2020)
```

### Property
```{r}
div(id = "body-section", state_bar_released_at_ped_property_2020)
```

### Public Order
```{r}
div(id = "body-section", state_bar_released_at_ped_publicorder_2020)
```

### Drug
```{r}
div(id = "body-section", state_bar_released_at_ped_drugs_2020)
```

### Other
```{r}
div(id = "body-section", state_bar_released_at_ped_other_2020)
```
:::

```{r}
div(id = "body-section", br(),
    div(id = "big-section-title", "Length of Stay"), br()
    )
```

::: {.panel-tabset}
### Overview
```{r}
div(id = "body-section", state_bar_sentence_overview_2020)
```

### Violent
```{r}
div(id = "body-section", state_bar_sentence_violent_2020)
```

### Property
```{r}
div(id = "body-section", state_bar_sentence_property_2020)
```

### Public Order
```{r}
div(id = "body-section", state_bar_sentence_publicorder_2020)
```

### Drug
```{r}
div(id = "body-section", state_bar_sentence_drugs_2020)
```

### Other
```{r}
div(id = "body-section", state_bar_sentence_other_2020)
```
:::
</center>
<br>



















## Disparities

<center>

```{r}
########################################

# Descriptive Statistics

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Descriptive Statistics"), br())
div(id = "body-section", "Race, ethincity, age, and gender", br(), br())
```

```{r}
########################################

# Relative Rate Index

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Relative Rate Index"), br(),
    div(id = "regular-text", "The relative rate index (RRI) provides a measure of disparity in outcomes for different racial and ethnic groups by comparing the rates at which individuals of different races and ethnic groups experience incarceration. The rate for each specific racial and ethnic group is computed by dividing the population of interest within that group by the total state population for that particular group."), br(),
    div(id = "regular-text", "To calculate the RRI, the rate of each outcome of interest for individuals who are Black non-Hispanic and Hispanic any race, is divided by the rate at which the outcome occurs for individuals who are White. A RRI of 1 indicates no disparity. RRIs greater than 1 indicate the outcome occurs more frequently for individuals who are Black or Hispanic, and an RRI less than 1 indicates an outcome occurs less frequently than for individuals who are White."),
br()
    )
```

::: {.panel-tabset}

### Sentence Length
```{r}
# div(id = "body-section", state_bar_rri_sentence_length_black)
# div(id = "body-section", state_bar_rri_sentence_length_hispanic)
div(
  #style = "display:flex; flex-direction: row; align-items: center; justify-items: center;",
  style = "display: grid; grid-template-columns: repeat(2, 1fr); justify-items: center;",
  state_bar_rri_sentence_length_black,
  state_bar_rri_sentence_length_hispanic
)
```

### Time Served
```{r}
div(id = "body-section", "Released before or on year of PED", br())
```

### Admission Type
```{r}
div(id = "body-section", "Parole return/revocation vs new crime", br())
```
:::


```{r}
########################################

# Predicted Probabilities

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Predicted Probabilities"), br())
div(id = "body-section", "Race, ethincity, age, gender, offense, and sentence length", br(), br())
```


```{r}
# div(id = "body-section", br(),
#     div(id = "big-section-title", "Racial and Ethnic Disparities"), br(),
#     div(id = "regular-text",      "Black people were X times less/more likely than White people to be released on or before their parole eligibility year."), br(), 
#     div(id = "regular-text",      "Hispanic people were X times less/more likely than White people to be released on or before their parole eligibility year."), br(), br(), 
# 
#     div(id = "regular-text",      "For every White person who is released after their parole eligibility year, X Black people are released after their parole eligibility year."), br(), 
#     div(id = "regular-text",      "For every White person who is released after their parole eligibility year, X Hispanic people are released after their parole eligibility year."), br(), 
#     div(id = "bar-graph",         state_time_between_release_ped_by_race), br(), br(),
#     div(id = "section-title",    "Predicted Probabilities of Being Released Within 1 Year of Parole Eligibility in 2020"), br(),
#      div(id = "reactable-table",
#          table_pp_by_race,
#          table_pp_by_sex,
#          table_pp_by_offgeneral,
#          table_pp_by_sentlgth)
#     )
```
</center>
<br>

::::























::: citation-section

Citations:   

Alper, Mariel E., et al. "Profiles in Parole Release and Revocation: Examining the Legal Framework in the United States." *Robina Institute of Criminal Law and Criminal Justice*, 13 May 2022, [robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states](robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states).     

Carson, Ann E. "Prisoners in 2020 – Statistical Tables." Bureau of Justice Statistics, 1 Dec. 2020, [bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables](bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables).    

United States. Bureau of Justice Statistics. Annual Parole Survey, 2018. Inter-university Consortium for Political and Social Research [distributor], 2021-10-28. [https://doi.org/10.3886/ICPSR38058.v1](https://doi.org/10.3886/ICPSR38058.v1). 
  
United States. Bureau of Justice Statistics. National Corrections Reporting Program, 1991-2020: Selected Variables. Inter-university Consortium for Political and Social Research [distributor], 2022-11-28. [https://doi.org/10.3886/ICPSR38492.v1](https://doi.org/10.3886/ICPSR38492.v1).     
:::

<br><br><br><br>

---
title: "`r state_for_report`"
---
