---
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

# load packages
library(shiny)
library(highcharter)
library(dplyr)
library(htmlwidgets)
library(dplyr)
library(htmltools)
library(reactable)
library(reactablefmtr)
library(htmltools)
library(tippy)
library(reactable)
library(tidyr)

# load data
source("prep/library.R")
source("prep/functions.R")
source("prep/dataframes.R")

state_for_report <- "Michigan"
```


```{r}
################################################################################

# Highlight Section at Top

################################################################################

# Create title for prison sentencing information
prison_sentencing_system_info <- HTML("Prison Sentencing<br>System&#9432;")

# Get prison sentencing information for state
prison_sentencing_system_name <- subset(robinainfo, state == state_for_report)$prison_sentencing_system

# Reformat text depending on prison sentencing information
prison_sentencing_system <- 
  case_when(
    prison_sentencing_system_name == "Extremely High Indeterminacy" ~ HTML("Extremely High<br>Indeterminacy"),
    prison_sentencing_system_name == "High Indeterminacy"           ~ HTML("High<br>Indeterminacy"),
    prison_sentencing_system_name == "Moderate Indeterminacy"       ~ HTML("Moderate<br>Indeterminacy"),
    prison_sentencing_system_name == "Low Indeterminacy"            ~ HTML("Low<br>Indeterminacy"),
    prison_sentencing_system_name == "Extremely Low Indeterminacy"  ~ HTML("Extremely Low<br>Indeterminacy"))

# Get prison sentencing information's definition
prison_sentencing_system_definition <- subset(robinadefinitions, term == prison_sentencing_system_name)$definition 

# Assemble all prison sentencing information for the state
prison_sentencing_system_tooltip <- 
  HTML(paste0("To compare the degrees of indeterminacy, or unpredictability
   of time served, of prison-sentencing systems of different jurisdictions,
   we use a qualitative ranking framework based on the Robina Institute's
   Prison-release Discretion And Prison Population Size Report.<br><br><b>", 
   prison_sentencing_system_name, "</b>: ",
   prison_sentencing_system_definition, "."))

# Make the information accessible
prison_sentencing_system_info <- paste0('<span aria-label="', prison_sentencing_system_tooltip, '" tabindex="0" role="button">', prison_sentencing_system_info, '</span>')

# Get number if people currently eligible and eligible in the future for parole
if (state_for_report %in% unique(parole_eligibility_table_2020$state)) {
  state_data <- parole_eligibility_table_2020 %>% filter(state == state_for_report)
} else {
  state_data <- "No Data"
}
num_people_current   <- sum(state_data$current_new_crime_count, na.rm = TRUE)
num_people_1_5_years <- sum(state_data$future_1_5_years_new_crime_count, na.rm = TRUE)






################################################################################

# Parole Eligibility Tab

################################################################################

###################
# Parole-Eligible Prison Population
###################

# Stacked bar chart showing the  proportion of parole eligibility types
if (state_for_report %in% names(all_stackedbar_pe_type_2020)) {
  state_stackedbar_pe_type_2020 <- 
    all_stackedbar_pe_type_2020[[state_for_report]] %>% 
    hc_size(height = 270) %>%
    hc_title(text = "Distribution of Parole Eligibility Statuses Among the Prison Population in 2020")
} else {
  state_stackedbar_pe_type_2020 <- "No Data"
}

# Save sentence about parole eligible population
if (state_for_report %in% names(all_sentence_parole_elgibility_population)) {
  state_sentence_parole_elgibility_population <- 
    all_sentence_parole_elgibility_population[[state_for_report]]
} else {
  state_sentence_parole_elgibility_population <- ""
}


###################
# How is Parole Eligibility Determined?
###################

# Get parole eligibility information by state
parole_eligibility_criteria <- subset(robinaparoleeligibility, state == state_for_report)$general_rules_of_release_eligibility


###################
# People who are Eligible for Parole
###################

# Save bar chart of people eligible for parole but still in prison by race and state
if (state_for_report %in% names(all_bar_parole_elgibility_race)) {
  state_bar_parole_elgibility_race <- 
    all_bar_parole_elgibility_race[[state_for_report]] %>% 
    hc_size(width = 600, height = 350) %>% 
    hc_title(text = "Distribution of Prison Population who are Parole-Eligible by Race in 2020") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_bar_parole_elgibility_race <- "No Data"
}

# Save sentence about PE by race and by state
if (state_for_report %in% names(all_sentence_parole_elgibility_race)) {
  state_sentence_parole_elgibility_race <- 
    all_sentence_parole_elgibility_race[[state_for_report]]
} else {
  state_sentence_parole_elgibility_race <- ""
}


###################
# Offense Types for the Parole-Eligible Prison Population
###################

# Save pie chart of most serious sentenced offense by state - new crime
if (state_for_report %in% names(all_pie_parole_elgibility_offense_new_crime)) {
  state_pie_parole_elgibility_offense_new_crime <- 
    all_pie_parole_elgibility_offense_new_crime[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Most Serious Sentenced Offense for Parole-Eligible People in Prison in 2020") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_pie_parole_elgibility_offense_new_crime <- "No Data"
}

# Save sentence about serious offense by state
if (state_for_report %in% names(all_sentence_parole_elgibility_offense)) {
  state_sentence_parole_elgibility_offense <- 
    all_sentence_parole_elgibility_offense[[state_for_report]]
} else {
  state_sentence_parole_elgibility_offense <- ""
}







################################################################################

# Prison Population Tab

################################################################################

# Stacked bar chart showing distribution of prison population by admission type
if (state_for_report %in% names(all_stackedbar_admtype_2020)) {
  state_stackedbar_admtype_2020 <-
    all_stackedbar_admtype_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "Distribution of the Prison Population by Admission Type")
} else {
  state_stackedbar_admtype_2020 <- "No Data"
}


###################
# Who is in Prison?
###################

# Stacked bar chart showing distribution of people by race
# New crime
if (state_for_report %in% names(all_stackedbar_new_crime_race_2020)) {
  state_stackedbar_new_crime_race_2020 <-
    all_stackedbar_new_crime_race_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "New Crime: Race/Ethincity")
} else {
  state_stackedbar_new_crime_race_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by sex
# New crime
if (state_for_report %in% names(all_stackedbar_new_crime_sex_2020)) {
  state_stackedbar_new_crime_sex_2020 <-
    all_stackedbar_new_crime_sex_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "New Crime: Gender")
} else {
  state_stackedbar_new_crime_sex_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by age
# New crime
if (state_for_report %in% names(all_stackedbar_new_crime_ageyrend_2020)) {
  state_stackedbar_new_crime_ageyrend_2020 <-
    all_stackedbar_new_crime_ageyrend_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "New Crime: Age")
} else {
  state_stackedbar_new_crime_ageyrend_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by race
# Parole return/revocation
if (state_for_report %in% names(all_stackedbar_parole_return_race_2020)) {
  state_stackedbar_parole_return_race_2020 <-
    all_stackedbar_parole_return_race_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "Parole return/revocation: Race/Ethincity")
} else {
  state_stackedbar_parole_return_race_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by sex
# Parole return/revocation
if (state_for_report %in% names(all_stackedbar_parole_return_sex_2020)) {
  state_stackedbar_parole_return_sex_2020 <-
    all_stackedbar_parole_return_sex_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "Parole return/revocation: Gender")
} else {
  state_stackedbar_parole_return_sex_2020 <- "No Data"
}

# Stacked bar chart showing distribution of people by age
# Parole return/revocation
if (state_for_report %in% names(all_stackedbar_parole_return_ageyrend_2020)) {
  state_stackedbar_parole_return_ageyrend_2020 <-
    all_stackedbar_parole_return_ageyrend_2020[[state_for_report]] %>%
    hc_size(height = 200) %>%
    hc_title(text = "Parole return/revocation: Age")
} else {
  state_stackedbar_parole_return_ageyrend_2020 <- "No Data"
}

###################
# Prison Population Trends
###################

# line chart showing the change in prison pop 
# and the change in releases to parole over time
if (state_for_report %in% names(all_line_pop_released_to_parole)) {
  state_line_pop_released_to_parole <- 
    all_line_pop_released_to_parole[[state_for_report]] %>% 
    hc_size(width = 600, height = 400) %>%
    hc_title(text = "Trends in Prison Population, Releases to Parole, and Parole Eligibility")
} else {
  state_line_pop_released_to_parole <- "No Data"
}






################################################################################

# Releases from Prison Tab

################################################################################

###################
# Max-Out vs. Early Releases
###################

# Save pie chart of whether people maxed our or didn't max out - new crime
if (state_for_report %in% names(all_pie_maxout_2020)) {
  state_pie_maxout_2020 <- 
    all_pie_maxout_2020[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Distribution of People who Maxed Out Their Sentences in 2020") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_pie_maxout_2020 <- "No Data"
}


###################
# Unconditional vs Conditional Releases
###################

# Save pie chart of whether people maxed our or didn't max out - new crime
if (state_for_report %in% names(all_pie_release_type_2020)) {
  state_pie_release_type_2020 <- 
    all_pie_release_type_2020[[state_for_report]] %>% 
    hc_size(width = 600, height = 300) %>% 
    hc_title(text = "Distribution of People Released Conditionally and Unconditionally in 2020") %>% 
    hc_subtitle(text = "(For Perople Incarcerated for a New Crime)")
} else {
  state_pie_release_type_2020 <- "No Data"
}


###################
# Release Timing by Offense Type
###################

# Save release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_2020)) {
  state_bar_released_at_ped_2020 <- 
    all_bar_released_at_ped_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/All Offenses)")
} else {
  state_bar_released_at_ped_2020 <- "No Data"
}

# Save other offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_violent_2020)) {
  state_bar_released_at_ped_violent_2020 <- 
    all_bar_released_at_ped_violent_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Violent Offense)")
} else {
  state_bar_released_at_ped_violent_2020 <- "No Data"
}

# Save publicorder offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_publicorder_2020)) {
  state_bar_released_at_ped_publicorder_2020 <- 
    all_bar_released_at_ped_publicorder_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Public Order Offense)")
} else {
  state_bar_released_at_ped_publicorder_2020 <- "No Data"
}

# Save property offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_property_2020)) {
  state_bar_released_at_ped_property_2020 <- 
    all_bar_released_at_ped_property_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Property Offense)")
} else {
  state_bar_released_at_ped_property_2020 <- "No Data"
}

# Save drug offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_drugs_2020)) {
  state_bar_released_at_ped_drugs_2020 <- 
    all_bar_released_at_ped_drugs_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Drug Offense)")
} else {
  state_bar_released_at_ped_drugs_2020 <- "No Data"
}

# Save other offense by release timing by state
if (state_for_report %in% names(all_bar_released_at_ped_other_2020)) {
  state_bar_released_at_ped_other_2020 <- 
    all_bar_released_at_ped_other_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Parole Eligibility Year vs. Actual Release Year in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Other Offense)")
} else {
  state_bar_released_at_ped_other_2020 <- "No Data"
}


###################
# Length of Stay by Offense Type
###################

# Save sentence duration by by state
if (state_for_report %in% names(all_bar_los_overview_2020)) {
  state_bar_los_overview_2020 <- 
    all_bar_los_overview_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/All Offenses)")
} else {
  state_bar_los_overview_2020 <- "No Data"
}

# Save sentence duration by offense type (violent) and by state
if (state_for_report %in% names(all_bar_los_violent_2020)) {
  state_bar_los_violent_2020 <- 
    all_bar_los_violent_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Violent Offense)")
} else {
  state_bar_los_violent_2020 <- "No Data"
}

# Save sentence duration by offense type (drugs) and by state
if (state_for_report %in% names(all_bar_los_drugs_2020)) {
  state_bar_los_drugs_2020 <- 
    all_bar_los_drugs_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Drug Offense)")
} else {
  state_bar_los_drugs_2020 <- "No Data"
}

# Save sentence duration by offense type (property) and by state
if (state_for_report %in% names(all_bar_los_property_2020)) {
  state_bar_los_property_2020 <- 
    all_bar_los_property_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Property Offense)")
} else {
  state_bar_los_property_2020 <- "No Data"
}

# Save sentence duration by offense type (public order) and by state
if (state_for_report %in% names(all_bar_los_publicorder_2020)) {
  state_bar_los_publicorder_2020 <- 
    all_bar_los_publicorder_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Public Order Offense)")
} else {
  state_bar_los_publicorder_2020 <- "No Data"
}

# Save sentence duration by offense type (other) and by state
if (state_for_report %in% names(all_bar_los_other_2020)) {
  state_bar_los_other_2020 <- 
    all_bar_los_other_2020[[state_for_report]] %>%
    hc_size(height = 100, width = 200) %>%
    hc_title(text = "Sentence Length Served for People Released in 2020") %>% 
    hc_subtitle(text = "(For People Incarcerated for New Crime/Other Offense)")
} else {
  state_bar_los_other_2020 <- "No Data"
}








################################################################################

# Disparities Tab

################################################################################

###################
# Descriptive Statistics
###################




###################
# Relative Rate Index (RRI)
###################

# Graph showing RRI for Black people
if (state_for_report %in% names(all_bar_rri_sentence_length_black)) {
  state_bar_rri_sentence_length_black <- 
    all_bar_rri_sentence_length_black[[state_for_report]] %>% 
    hc_size(width = 300, height = 300) 
} else {
  state_bar_rri_sentence_length_black <- "No Data"
}

# Graph showing RRI for Hispanic people
if (state_for_report %in% names(all_bar_rri_sentence_length_hispanic)) {
  state_bar_rri_sentence_length_hispanic <- 
    all_bar_rri_sentence_length_hispanic[[state_for_report]] %>% 
    hc_size(width = 300, height = 300) 
} else {
  state_bar_rri_sentence_length_hispanic <- "No Data"
}
```

::: narrow-section
```{r}
########################################

# Highlights Section

########################################

htmltools::div(
  id = "grid-container",
  style = "display: grid; grid-template-columns: repeat(3, 1fr); justify-items: center;",
  div(id = "highlight-text", formattable::comma(num_people_current, digits = 0)),
  div(id = "highlight-text", formattable::comma(num_people_1_5_years, digits = 0)),
  div(style ="display: block; 
              color: #ff6400; 
              font-size: 1.25em; 
              font-family: Graphik; 
              font-weight: bold; 
              text-align: center;",
      prison_sentencing_system),
  div(id = "bold-text", HTML("Currently Eligible<br>for Parole")),
  div(id = "bold-text", HTML("Eligible for Parole<br>in the next 1-5 years")),
  div(id = "bold-text", tippy(prison_sentencing_system_info,
                            placement = "bottom",
                            flip = TRUE,
                            theme = "light",
                            tooltip = prison_sentencing_system_tooltip))
)
```
:::
<br><br>














:::: {.panel-tabset .nav-pills}

## Prison Population
<center>
```{r}
########################################

# Prison Population by Admission Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "New Court Commits vs. Parole Revocations"), br(),
    div(id = "regular-text", "While the public often envisions a prison population primarily made up of individuals serving long sentences for serious crimes, the reality is far more nuanced. In particular, the distinction between new court commits—those incarcerated for the first time on a new charge—and parole revocations—those returned to prison for violating the conditions of their parole—is a critical one. Understanding the proportion of these two groups within the prison population can provide valuable insights into the effectiveness of the state's criminal justice policies, including sentencing and parole practices."), br(),
    div(id = "visualization", state_stackedbar_admtype_2020), br()
    )
```

```{r}
########################################

# Prison Population Trends

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Prison Population Trends"), br(),
    div(id = "visualization", state_line_pop_released_to_parole), br()
    )
```

```{r}
########################################

# Who's in Prison?

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Who's in Prison?"), br(),
    div(id = "regular-text", "[Race/Ethnicity][Gender] in who are between [Age Range] years old made up the most people in prison in 2020"), br(),
    )
```
::: {.panel-tabset}
### Race
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_new_crime_race_2020), br(),
    div(id = "visualization", state_stackedbar_parole_return_race_2020), br()
    )
```

### Gender
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_new_crime_sex_2020), br(),
    div(id = "visualization", state_stackedbar_parole_return_sex_2020), br()
    )
```

### Age
```{r}
div(id = "body-section", br(),
    div(id = "visualization", state_stackedbar_new_crime_ageyrend_2020), br(),
    div(id = "visualization", state_stackedbar_parole_return_ageyrend_2020), br()
    )
```

### Offense Types
```{r}
div(id = "body-section", br(),
    div(id = "visualization", "Visualization by FBI violent index, prop, drug, public order"), br()
    )
```

### Sentence Length
```{r}
div(id = "body-section", br(),
    div(id = "visualization", "TBD"), br()
    )
```
:::
</center>
<br>













## Parole Eligibility

<center>
```{r}

########################################

# How is Parole Eligibility Determined?

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "How is Parole Eligibility Determined?"), br(),
    div(id = "regular-text",       parole_eligibility_criteria),            br(), br()
)

########################################

# Parole-Eligible Prison Population

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Parole-Eligible Prison Population"),           br(),
    div(id = "regular-text",       state_sentence_parole_elgibility_population),  br(), 
    div(id = "bar-chart",          state_stackedbar_pe_type_2020),                br()
)

########################################

# People who are Eligible for Parole

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "People who are Eligible for Parole"), br()
    )
```
::: {.panel-tabset}
### Race
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  state_sentence_parole_elgibility_race), br(),
    div(id = "visualization", state_bar_parole_elgibility_race)
    )
```

### Gender

### Age
:::

```{r}
########################################

# Sentence Lengths

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Sentence Lengths"), br(),
    div(id = "regular-text", "TBD"), br()
    )
```

```{r}
########################################

# Offense Types

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Offense Types"), br(),
    div(id = "regular-text", state_sentence_parole_elgibility_offense), br(),
    state_pie_parole_elgibility_offense_new_crime)
```
</center>
<br>














## Releases

<center>
```{r}
########################################

# Max-Out vs. Early Releases

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Max-Out vs. Early Releases"), br(),
    div(id = "regular-text",  "In 2020, there were X number of people who maxed out their sentence lengths, meaning they were released on or after their mandatory release year, accounting for X% of releases from prison."), br(),
    state_pie_maxout_2020, br(), 
)

########################################

# Unconditional vs Conditional Releases

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Unconditional vs Conditional Releases"), br(),
    div(id = "regular-text",  "In 2020, there were X number of people who were released conditionally from prison, meaning they were released from incarceration under specific conditions and supervision and often before they served their full sentence, accounting for X% of releases from prison."), br(),
    state_pie_release_type_2020, br()
)


########################################

# Release Timing by Offense Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Release Timing by Offense Type"), br()
)
```

::: {.panel-tabset}
### Overview
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  "In 2020, for people in prison for a new crime, X number of people (X%) were released from prison on their parole eligibility year."), br(),
    state_bar_released_at_ped_2020)
```

### Violent
```{r}
div(id = "body-section", state_bar_released_at_ped_violent_2020)
```

### Property
```{r}
div(id = "body-section", state_bar_released_at_ped_property_2020)
```

### Public Order
```{r}
div(id = "body-section", state_bar_released_at_ped_publicorder_2020)
```

### Drug
```{r}
div(id = "body-section", state_bar_released_at_ped_drugs_2020)
```

### Other
```{r}
div(id = "body-section", state_bar_released_at_ped_other_2020)
```
:::

```{r}
########################################

# Time Served by Offense Type

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Time Served by Offense Type"), br()
    )
```

::: {.panel-tabset}
### Overview
```{r}
div(id = "body-section", br(),
    div(id = "regular-text",  "In 2020, for people in prison for a new crime, X number of people (X%) served their full sentence length.", br(), br()),
    state_bar_los_overview_2020)
```

### Violent
```{r}
div(id = "body-section", state_bar_los_violent_2020)
```

### Property
```{r}
div(id = "body-section", state_bar_los_property_2020)
```

### Public Order
```{r}
div(id = "body-section", state_bar_los_publicorder_2020)
```

### Drug
```{r}
div(id = "body-section", state_bar_los_drugs_2020)
```

### Other
```{r}
div(id = "body-section", state_bar_los_other_2020)
```
:::
</center>
<br>



















## Disparities

<center>

```{r}
########################################

# Descriptive Statistics

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Descriptive Statistics"), br())
div(id = "body-section", "Race, ethincity, age, and gender", br(), br())
```

```{r}
########################################

# Relative Rate Index

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Relative Rate Index"), br(),
    div(id = "regular-text", "The relative rate index (RRI) provides a measure of disparity in outcomes for different racial and ethnic groups by comparing the rates at which individuals of different races and ethnic groups experience incarceration. The rate for each specific racial and ethnic group is computed by dividing the population of interest within that group by the total state population for that particular group."), br(),
    div(id = "regular-text", "To calculate the RRI, the rate of each outcome of interest for individuals who are Black non-Hispanic and Hispanic any race, is divided by the rate at which the outcome occurs for individuals who are White. A RRI of 1 indicates no disparity. RRIs greater than 1 indicate the outcome occurs more frequently for individuals who are Black or Hispanic, and an RRI less than 1 indicates an outcome occurs less frequently than for individuals who are White."),
br()
    )
```

::: {.panel-tabset}

### First Time Offenses
```{r}
div(id = "body-section", 
    div(id = "regular-text", "Compare parole eligibility rates or rates of release to parole for Black, White, and Hispanic individuals with similar first time offenses. For example, are Black people less likely to be released to parole on their parole eligibility date for drug-related convictions compared to White people?", br())
)
```

### Time Served
```{r}
div(id = "body-section", 
    div(id = "regular-text", "Compare time served for Black, White, and Hispanic individuals with similar sentence lengths and offenses.", br())
)
```

### Sentence Length (not using)
```{r}
div(
  style = "display: grid; grid-template-columns: repeat(2, 1fr); justify-items: center;",
  state_bar_rri_sentence_length_black,
  state_bar_rri_sentence_length_hispanic
)
```
:::


```{r}
########################################

# Predicted Probabilities

########################################
div(id = "body-section", br(),
    div(id = "big-section-title", "Predicted Probabilities"), br())
div(id = "body-section", "Race, ethincity, age, gender, offense, and sentence length", br(), br())
```


```{r}
# div(id = "body-section", br(),
#     div(id = "big-section-title", "Racial and Ethnic Disparities"), br(),
#     div(id = "regular-text",      "Black people were X times less/more likely than White people to be released on or before their parole eligibility year."), br(), 
#     div(id = "regular-text",      "Hispanic people were X times less/more likely than White people to be released on or before their parole eligibility year."), br(), br(), 
# 
#     div(id = "regular-text",      "For every White person who is released after their parole eligibility year, X Black people are released after their parole eligibility year."), br(), 
#     div(id = "regular-text",      "For every White person who is released after their parole eligibility year, X Hispanic people are released after their parole eligibility year."), br(), 
#     div(id = "bar-graph",         state_time_between_release_ped_by_race), br(), br(),
#     div(id = "section-title",    "Predicted Probabilities of Being Released Within 1 Year of Parole Eligibility in 2020"), br(),
#      div(id = "reactable-table",
#          table_pp_by_race,
#          table_pp_by_sex,
#          table_pp_by_offgeneral,
#          table_pp_by_sentlgth)
#     )
```
</center>
<br>

::::























::: citation-section

Citations:   

Alper, Mariel E., et al. "Profiles in Parole Release and Revocation: Examining the Legal Framework in the United States." *Robina Institute of Criminal Law and Criminal Justice*, 13 May 2022, [robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states](robinainstitute.umn.edu/publications/profiles-parole-release-and-revocation-examining-legal-framework-united-states).     

Carson, Ann E. "Prisoners in 2020 – Statistical Tables." Bureau of Justice Statistics, 1 Dec. 2020, [bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables](bjs.ojp.gov/library/publications/prisoners-2020-statistical-tables).    

United States. Bureau of Justice Statistics. Annual Parole Survey, 2018. Inter-university Consortium for Political and Social Research [distributor], 2021-10-28. [https://doi.org/10.3886/ICPSR38058.v1](https://doi.org/10.3886/ICPSR38058.v1). 
  
United States. Bureau of Justice Statistics. National Corrections Reporting Program, 1991-2020: Selected Variables. Inter-university Consortium for Political and Social Research [distributor], 2022-11-28. [https://doi.org/10.3886/ICPSR38492.v1](https://doi.org/10.3886/ICPSR38492.v1).     
:::

<br><br><br>

---
title: "`r state_for_report`"
---
