---
title: "Missing Data"
format:
  html:
    css: styles.css
    embed-resources: TRUE
    code-fold: true
    page-layout: full
    fig_caption: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png"
)

source("prep/library.R")
source("prep/functions.R")

# load data
load(file = paste0(sp_data_path, "/data/analysis/all_sentlgth_ped.rds"))
load(file = paste0(sp_data_path, "/data/analysis/all_sentlgth_125_ped.rds"))
# load(file = paste0(sp_data_path, "/data/analysis/all_offgeneral_ped.rds"))
# load(file = paste0(sp_data_path, "/data/analysis/all_offdetail_ped.rds"))
# load(file = paste0(sp_data_path, "/data/analysis/all_sentlgth_offgeneral.rds"))
# load(file = paste0(sp_data_path, "/data/analysis/all_offgeneral_sentlgth.rds"))
```

**Questions**

1. **Depending on the state, can we assume that people in prison with missing parole eligibility data who have a sentence length of "Life, LWOP, Life plus additional years, Death" are not eligible for parole? Does offense type matter?** For example, in the state of Georgia, 30% of people have missing parole eligibility data. Of these people, 70% had a violent offense and 20% had a Life/LWOP/Death sentence length. Of the people with a Life/LWOP/Death sentence length and missing parole eligibility data, 97% had a violent offense.   
<br>
2. **Depending on the state, can we assume that someone is not eligible for parole because they have a short sentence? Is time served so brief that the parole process might not be deemed necessary?** For example, in Georgia, 83% of people with a sentence length of <1 year are missing parole eligibility information.   
<br>
3. **Do we need to take into account the prison sentencing system of the state to determine who is potentially eligible or not eligible for parole?** For example, in Georgia,  "While the grant of parole is discretionary, consideration for parole is often not; most inmates are entitled to automatic consideration for parole. In general, this occurs after serving 1/3 of the total sentence. However, at minimum, a person serving a misdemeanor sentence must serve 6 months and a person serving a felony sentence must serve 9 months. Inmates serving 21 years or more, but not a life sentence, become eligible after serving a 7 year term. In addition, inmates convicted of certain serious crimes (e.g. aggravated assault, incest, drug trafficking) are not eligible for parole until they have served 7 years or 1/3 of the prison term, whichever comes first."  
<br>
[Word document](https://csgorg.sharepoint.com/:w:/s/Team-JC-Research/EZCeIZ0UTgNCu9aCZ4eXqpYBsHeh_ETS4EM3a-nOJmB5Cg?e=BXUemM) in Sharepoint for discussing missingness.  

```{r}
################################################################################

# Missingness by sentence length and parole eligibility status

# Obtained from NCRP year end population

################################################################################

# reactable table
table1 <-
  reactable(all_sentlgth_125_ped,
            style = reactable_style,
            theme = reactable_theme,
            defaultColDef = colDef(format = colFormat(separators = TRUE),
                                   align = "right",
                                   minWidth = 70),
            compact = TRUE,
            defaultPageSize = 50,
            columnGroups = list(
              colGroup(name = "Number of People",
                       columns = c("current",
                                   "future_1_5_years",
                                   "future_6_years",
                                   "missing")),
              colGroup(name = "Percentage of the Prison Population",
                       columns = c("current_percent",
                                   "future_1_5_years_percent",
                                   "future_6_years_percent",
                                   "missing_percent"))),
            columns = list(
              state         = colDef(name = "State",
                                     minWidth = 120,
                                     align = "left",
                                     style = list(fontWeight = "bold")
                                     ),
              sentlgth_category      = colDef(name = "Sentence Length",
                                       style = list(fontWeight = "bold"),
                                       minWidth = 120,
                                       align = "left"),

              current          = colDef(name = "Current"),
              future_1_5_years = colDef(name = "Future 1-5 Years"),
              future_6_years   = colDef(name = "Future 6+ Years"),
              missing          = colDef(name = "Missing"),

              current_percent          = colDef(name = "Current",
                                                format = colFormat(percent = TRUE, digits = 0)),
              future_1_5_years_percent = colDef(name = "Future 1-5 Years",
                                                format = colFormat(percent = TRUE, digits = 0)),
              future_6_years_percent   = colDef(name = "Future 6+ Years",
                                                format = colFormat(percent = TRUE, digits = 0)),
              missing_percent          = colDef(name = "Missing",
                                                style = list(fontWeight = "bold"),
                                                format = colFormat(percent = TRUE, digits = 0)),

              total           = colDef(name = "Total",
                                       style = list(borderRight = "1px solid #d3d3d3"))
            )
  )


# reactable table
table2 <-
  reactable(all_sentlgth_ped,
            style = reactable_style,
            theme = reactable_theme,
            defaultColDef = colDef(format = colFormat(separators = TRUE),
                                   align = "right",
                                   minWidth = 70),
            groupBy = "state",
            compact = TRUE,
            filterable = TRUE,
            defaultExpanded = TRUE,
            defaultPageSize = 3,
            columnGroups = list(
              colGroup(name = "Number of People",
                       columns = c("current",
                                   "future_1_5_years",
                                   "future_6_years",
                                   "missing")),
              colGroup(name = "Percentage of the Prison Population",
                       columns = c("current_percent",
                                   "future_1_5_years_percent",
                                   "future_6_years_percent",
                                   "missing_percent"))),
            columns = list(
              state         = colDef(name = "State",
                                     minWidth = 120,
                                     align = "left",
                                     style = list(fontWeight = "bold")),
              sentlgth      = colDef(name = "Sentence Length",
                                       style = list(fontWeight = "bold"),
                                       minWidth = 120,
                                       align = "left"),

              current          = colDef(name = "Current"),
              future_1_5_years = colDef(name = "Future 1-5 Years"),
              future_6_years   = colDef(name = "Future 6+ Years"),
              missing          = colDef(name = "Missing"),

              current_percent          = colDef(name = "Current",
                                                format = colFormat(percent = TRUE, digits = 0)),
              future_1_5_years_percent = colDef(name = "Future 1-5 Years",
                                                format = colFormat(percent = TRUE, digits = 0)),
              future_6_years_percent   = colDef(name = "Future 6+ Years",
                                                format = colFormat(percent = TRUE, digits = 0)),
              missing_percent          = colDef(name = "Missing",
                                                style = list(fontWeight = "bold"),
                                                format = colFormat(percent = TRUE, digits = 0)),

              total           = colDef(name = "Total",
                                       style = list(borderRight = "1px solid #d3d3d3"))
            )
  )
```

<br>






**Parole Eligibility Status by Sentence Length**

Data: NCRP year-end population  

This table includes people in prison for a new court commitment or parole return/revocation and with sentence lengths between 1-24.9 years.  

```{r}
table1
```
<br><br>

Data: NCRP year-end population  

This table includes people in prison for a new court commitment or parole return/revocation and with any sentence length.  

```{r}
table2
```

