---
title: "Untitled"
editor: visual
---

# Disparities

## Incarceration Rates

`r paste(state_sentence_rri, collapse = ", ")`

```{r state_hc_waffle_rri_demographics}
# div(id = "visualization", state_hc_waffle_rri_black)
# div(id = "visualization", state_hc_waffle_rri_white)
# div(id = "visualization", state_hc_waffle_rri_hispanic)
# div(id = "visualization", state_hc_waffle_rri_other)
```

<br>

::: divider
:::

## Years Spent in Prison After Parole Eligibility

::: {.panel-tabset .custom-tabs}
### Race and Ethnicity

```{r state_parole_release_disparities, results='asis'}
df1 <- all_parole_release_disparities |>
  filter(state == state_for_report) |> 
  filter(!is.na(race)) |> 
  select(time_between_ped_release, race) |> 
  group_by(race) |> 
  summarise(total_years = sum(time_between_ped_release, na.rm = TRUE))

# Check if all total_years are NA
if (!all(is.na(df1$total_years))) {
  
  message <- HTML(paste0("In ", select_year, ", people in prison past their parole eligibility year collectively spent a total of ",
                   comma(sum(df1$total_years, na.rm = TRUE)), " years behind bars instead of supervised in the community on parole."))
  
  cat(message)
  
  br()
  html_output <- div(
    id = "grid-container",
    style = "display: grid; grid-template-columns: repeat(4, 1fr); grid-row-gap: 0px; grid-column-gap: 0px;",
    div(style = "color: #49a7a1; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Black, non-Hispanic"]) ),
    div(style = "color: #55b4e5; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "White, non-Hispanic"]) ),
    div(style = "color: #d97d68; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Hispanic, any race"]) ),
    div(style = "color: #decf64; text-align: center; font-size: 24px; font-weight: bold;", comma(df1$total_years[df1$race == "Other race(s), non-Hispanic"]) ),
    div(style = "color: #49a7a1; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #55b4e5; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #d97d68; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "color: #decf64; text-align: center; font-size: 20px; font-weight: bold;", "Years"),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Black People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("White People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Hispanic People")),
    div(style = "text-align: center; font-size: 14px; font-weight: bold;", HTML("Other Race(s), Non-Hispanic People"))
  )
  
  cat(as.character("<br><br>"))
  cat(as.character(html_output))
  br()
} else {
  cat("")
}
```

```{r state_scatter_race_ped_release}
# div(id = "visualization", state_scatter_race_ped_release)
```

### Gender

In progress. Same type of graph as race and ethnicity tab.
:::

<br>

::: divider
:::

## Timing of Release After Parole Eligibility

::: {.panel-tabset .custom-tabs}
### Race and Ethnicity

The timing of release for individuals after their parole eligibility year varied by offense type and demographic group, with many remaining in prison for several years beyond their eligibility date.

```{r state_bubble_race_ped_release}
# div(id = "visualization", state_bubble_race_ped_release)
```

### Gender

In progress. Same type of graph as race and ethnicity tab.
:::

<br>

::: divider
:::

## Time Served

::: {.panel-tabset .custom-tabs}
### Race and Ethnicity

```{r}
# section_text <- "X individuals faced the longest average time served in prison in 2020. Similarly, Y, individuals also experienced notably longer prison stays compared to their Z counterparts, who had the shortest average time served."
# div(
#   id = "grid-container",
#   style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-row-gap: 0px; grid-column-gap: 20px;",
#   section_text,
#   div(id = "visualization", state_lollipop_los_race)
# )
# br()
```

`r paste(state_sentence_los_race, collapse = ", ")`

```{r state_lollipop_los_race}
div(id = "visualization", style = "margin: 0 auto; text-align: center;width: 400px", state_lollipop_los_race)
# div(id = "visualization", style = "margin: 0 auto; text-align: center;", state_lollipop_los_race)
```

`r paste(state_sentence_los_race_offense, collapse = ", ")`

```{r state_scatter_los_race_offense}
div(id = "visualization", state_scatter_los_race_offense)
```

### Gender

In progress. Same type of graph as race and ethnicity tab.
:::

<br> <br>
