---
title: "Untitled"
editor: visual
---

# Prison Population

## Prison Population Trends

`r paste(state_sentence_population, collapse = ", ")`

```{r state_line_population_by_year}
div(id = "visualization", state_line_population_by_year)
```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners Series</a>

<br>

::: divider
:::

## Demographics

`r paste(state_sentence_population_demographics, collapse = ", ")`

```{r state_population_demographics}

```

<a href="https://bjs.ojp.gov/library/publications/prisoners-2022-statistical-tables" style="font-size: 0.7em; color: gray;">Source: BJS Prisoners 2022 - Race and Ethnicity</a>

<br>

::: divider
:::

## Criminal Offenses

`r paste(state_sentence_population_fbi_index, collapse = ", ")`

```{r state_bar_population_fbi_index}
div(id = "visualization", state_bar_population_fbi_index)
```

<br>

::: divider
:::

## Sentence Lengths

`r paste(state_sentence_population_sentlgth, collapse = ", ")`

```{r state_bar_population_sentlgth}
div(id = "visualization", state_bar_population_sentlgth)
```

<br>
